### 2024-02-29 glossaries.tex

\newacronym{af}{AF}{Amplify-and-Forward}
\newacronym{ambc}{AmBC}{Ambient Backscatter Communication}
\newacronym{am}{AM}{Arithmetic Mean}
\newacronym{ao}{AO}{Alternating Optimization}
\newacronym{ap}{AP}{Access Point}
\newacronym{awgn}{AWGN}{Additive White Gaussian Noise}
\newacronym{bbc}{BBC}{Bistatic \glsentryshort{bc}}
\newacronym{bcd}{BCD}{Block Coordinate Descent}
\newacronym{bc}{BackCom}{Backscatter Communication}
\newacronym{bd}{BD}{Beyond-Diagonal}
\newacronym{ber}{BER}{Bit Error Rate}
\newacronym{bibo}{BIBO}{Binary-Input Binary-Output}
\newacronym{bls}{BLS}{Backtracking Line Search}
\newacronym{bpcu}{\si{bpcu}}{bits per channel use}
\newacronym{bpsphz}{\si{bps/Hz}}{bits per second per Hertz}
\newacronym{clt}{CLT}{Central Limit Theorem}
\newacronym{cp}{CP}{Canonical Polyadic}
\newacronym{cr}{CR}{Cognitive Radio}
\newacronym{cscg}{CSCG}{Circularly Symmetric Complex Gaussian}
\newacronym{csit}{CSIT}{\gls{csi} at the Transmitter}
\newacronym{csi}{CSI}{Channel State Information}
\newacronym{cw}{CW}{Continuous Waveform}
\newacronym{dcmc}{DCMC}{Discrete-input Continuous-output Memoryless Channel}
\newacronym{dc}{DC}{Direct Current}
\newacronym{df}{DF}{Decode-and-Forward}
\newacronym{dmc}{DMC}{Discrete Memoryless Channel}
\newacronym{dmmac}{DMMAC}{Discrete Memoryless Multiple Access Channel}
\newacronym{dmtc}{DMTC}{Discrete Memoryless Thresholding Channel}
\newacronym{dof}{DoF}{Degree of Freedom}
\newacronym{dp}{DP}{Dynamic Programming}
\newacronym{eirp}{EIRP}{Effective Isotropic Radiated Power}
\newacronym{fdma}{FDMA}{Frequency-Division Multiple Access}
\newacronym{fpga}{FPGA}{Field-Programmable Gate Array}
\newacronym{gm}{GM}{Geomtric Mean}
\newacronym{gp}{GP}{Geometric Programming}
\newacronym{ic}{IC}{Interference Channel}
\newacronym{iid}{i.i.d.}{independent and identically distributed}
\newacronym{ioe}{IoE}{Internet of Everything}
\newacronym{iot}{IoT}{Internet of Things}
\newacronym{kkt}{KKT}{Karush-Kuhn-Tucker}
\newacronym{los}{LoS}{Line-of-Sight}
\newacronym{m2m}{M2M}{Machine-to-Machine}
\newacronym{mac}{MAC}{Multiple Access Channel}
\newacronym{mbc}{MBC}{Monostatic \glsentryshort{bc}}
\newacronym{mc}{MC}{Multiplication Coding}
\newacronym{mimo}{MIMO}{Multiple-Input Multiple-Output}
\newacronym{miso}{MISO}{Multiple-Input Single-Output}
\newacronym{ml}{ML}{Maximum-Likelihood}
\newacronym{mmse}{MMSE}{Minimum Mean-Square-Error}
\newacronym{mrc}{MRC}{Maximal Ratio Combining}
\newacronym{mrt}{MRT}{Maximum Ratio Transmission}
\newacronym{mse}{MSE}{Mean-Square Error}
\newacronym{nlos}{NLoS}{Non-\gls{los}}
\newacronym{noma}{NOMA}{Non-Orthogonal Multiple Access}
\newacronym{ofdm}{OFDM}{Orthogonal Frequency-Division Multiplexing}
\newacronym{pc}{PC}{Point-to-point Channel}
\newacronym{pdf}{PDF}{Probability Density Function}
\newacronym{pga}{PGA}{Projected Gradient Ascent}
\newacronym{pin}{PIN}{Positive Intrinsic Negative}
\newacronym{psk}{PSK}{Phase Shift Keying}
\newacronym{ps}{PS}{Power Splitting}
\newacronym{qam}{QAM}{Quadrature Amplitude Modulation}
\newacronym{qos}{QoS}{Quality of Service}
\newacronym{r-e}{R-E}{Rate-Energy}
\newacronym{rcg}{RCG}{Riemannian Conjugate Gradient}
\newacronym{rfid}{RFID}{\gls{rf} Identification}
\newacronym{rf}{RF}{Radio-Frequency}
\newacronym{ris}{RIS}{Reconfigurable Intelligent Surface}
\newacronym{sca}{SCA}{Successive Convex Approximation}
\newacronym{sc}{SC}{Superposition Coding}
\newacronym{sdma}{SDMA}{Space-Division Multiple Access}
\newacronym{sdp}{SDP}{Semi-Definite Programming}
\newacronym{sdr}{SDR}{Semi-Definite Relaxation}
\newacronym{sic}{SIC}{Successive Interference Cancellation}
\newacronym{simo}{SIMO}{Single-Input Multiple-Output}
\newacronym{sinr}{SINR}{Signal-to-Interference-plus-Noise Ratio}
\newacronym{siso}{SISO}{Single-Input Single-Output}
\newacronym{smawk}{SMAWK}{Shor-Moran-Aggarwal-Wilber-Klawe}
\newacronym{smf}{SMF}{Scaled Matched Filter}
\newacronym{snr}{SNR}{Signal-to-Noise Ratio}
\newacronym{sr}{SR}{Symbiotic Radio}
\newacronym{svd}{SVD}{Singular Value Decomposition}
\newacronym{swipt}{SWIPT}{Simultaneous Wireless Information and Power Transfer}
\newacronym{tdma}{TDMA}{Time-Division Multiple Access}
\newacronym{ts}{TS}{Time Switching}
\newacronym{ue}{UE}{User Equipment}
\newacronym{wf}{WF}{Water-Filling}
\newacronym{wit}{WIT}{Wireless Information Transfer}
\newacronym{wpcn}{WPCN}{Wireless Powered Communication Network}
\newacronym{wpt}{WPT}{Wireless Power Transfer}
\newacronym{zf}{ZF}{Zero-Forcing}

% \nomenclature[g-pi]{$\pi$}{ $\simeq 3.14\ldots$}
% \nomenclature[c-pi]{$\pi$}{$\simeq 3.14\ldots$}
% \nomenclature[c-pi]{$\pi$}{Archimedes' constant $\simeq 3.1415\ldots$}

\nomenclature[e-1]{$a,A$}{Scalar}
\nomenclature[e-2]{$\mathbf{a}$}{Column vector}
\nomenclature[e-3]{$\mathbf{A}$}{Matrix}
\nomenclature[e-3]{$\mathcal{A}$}{Finite set}
\nomenclature[e-4]{$\mathbf{0}$}{Zero matrix}
\nomenclature[e-5]{$\mathbf{1}$}{One matrix}
\nomenclature[e-6]{$\mathbf{I}$}{Identity matrix}

\nomenclature[n-1]{$\mathbb{R}$}{Real numbers}
\nomenclature[n-2]{$\mathbb{R}_+$}{Real nonnegative numbers}
\nomenclature[n-3]{$\mathbb{C}$}{Complex numbers}
\nomenclature[n-4]{$\mathbb{I}$}{Probability domain $[0,1]$}
\nomenclature[n-5]{$\mathbb{H}_+^{n \times n}$}{Positive semi-definite matrices of dimension $n \times n$}
\nomenclature[n-6]{$\mathbb{U}^{n \times n}$}{Unitary matrices of dimension $n \times n$}

\nomenclature[c-j]{$\jmath$}{The imaginary unit $= \sqrt{-1}$}
\nomenclature[c-e]{$e$}{Euler's number $\simeq 2.71828\ldots$}
\nomenclature[c-pi]{$\pi$}{Archimedes' constant $\simeq 3.14159\ldots$}

\nomenclature[o-a]{$(\cdot)^*$}{Complex conjugate}
\nomenclature[o-b]{$(\cdot)^\mathsf{T}$}{Transpose}
\nomenclature[o-c]{$(\cdot)^\mathsf{H}$}{Hermitian (conjugate transpose)}
\nomenclature[o-d]{$(\cdot)^\dagger$}{Moore-Penrose inverse}
\nomenclature[o-e]{$(\cdot)^+$}{Ramp functio $\max(0, \cdot)$}
\nomenclature[o-f]{$\lvert \cdot \rvert$}{Absolute value of a complex number}
\nomenclature[o-g]{$\lVert \cdot \rVert$}{Euclidean norm of a vector}
\nomenclature[o-h]{$\lVert \cdot \rVert _ \mathrm{F}$}{Frobenius norm of a matrix}
\nomenclature[o-i]{$\mathrm{arg}(\cdot)$}{Argument of a complex number}
\nomenclature[o-j]{$\mathrm{card}(\cdot)$}{Cardinality of a finite set}
\nomenclature[o-k]{$\log(\cdot)$}{Natural logarithm of a real number}
\nomenclature[o-l]{$\exp(\cdot)$}{Exponential of a scalar or square matrix}
\nomenclature[o-m]{$\mathrm{tr}(\cdot)$}{Trace of a square matrix}
\nomenclature[o-n]{$\mathrm{det}(\cdot)$}{Determinant of a square matrix}
\nomenclature[o-o]{$\mathrm{sv}(\cdot)$}{Singular values sorted from largest to smallest}
\nomenclature[o-p]{$\mathrm{diag}(\cdot)$}{Constructs a square matrix with inputs on the main diagonal}
\nomenclature[o-q]{$\mathrm{diag}^{-1}(\cdot)$}{Retrieves the main diagonal of a square matrix}
\nomenclature[o-r]{$\Re(\cdot)$}{Retrieves the real part of a complex number}
\nomenclature[o-s]{$\Im(\cdot)$}{Retrieves the imaginary part of a complex number}
\nomenclature[o-t]{$\mathbb{E}(\cdot)$}{Expectation operator}
\nomenclature[o-u]{$\mathbb{A}(\cdot)$}{Extracts the \glsdesc{dc} component of a signal}
\nomenclature[o-v]{$\odot$}{Hadamard product}
\nomenclature[o-w]{$\otimes$}{Kronecker product}
\nomenclature[o-x]{$(\cdot)_{[x:y]}$}{Shortcut for $(\cdot)_x,(\cdot)_{x+1},\ldots,(\cdot)_y$}

\nomenclature[r-a]{$\sim$}{Follows a distribution}
\nomenclature[r-b]{$\mathcal{CN}(\mathbf{0}, \mathbf{\Sigma})$}{Multivariate \glsdesc{cscg} with covariance $\mathbf{\Sigma}$}

\nomenclature[s-d]{$(\cdot)_\text{D}$}{Direct}
\nomenclature[s-f]{$(\cdot)_\text{F}$}{Forward}
\nomenclature[s-b]{$(\cdot)_\text{B}$}{Backward}
\nomenclature[s-i]{$(\cdot)_\text{I}$}{Information}
\nomenclature[s-p]{$(\cdot)_\text{P}$}{Power}

\nomenclature[u-a]{$(\cdot)^{(r)}$}{$r$-th iterated value}
\nomenclature[u-b]{$(\cdot)^\star$}{Stationary point}