### 2024-03-04 content/chapter_5.tex

%!TEX root = ../thesis.tex

\graphicspath{{assets/chapter_5/}}

\chapter{\glsfmtshort{mimo} Channel Shaping using \glsfmtshort{ris}: From Diagonal to Beyond}

\begin{section}{Introduction}
	% The quest for reliable, high-speed, and ubiquitous wireless connectivity has been long-standing since Marconi's illuminating radio in 1895.
	% Great successes have been made at transmitter and receiver sides over the past century, and the society is unprecedentedly close to the Shannon limit \cite{Shannon1948}.
	Today we are witnessing a paradigm shift from connectivity to intelligence, where the wireless environment is no longer a chaotic medium but a conscious agent that serves on demand.
	This is empowered by the recent advances in \gls{ris}, a real-time programmable metasurface of numerous non-resonant sub-wavelength scattering elements.
	It can manipulate the amplitude, phase, frequency, and polarization of the scattered waves \cite{Basar2019} with a higher energy efficiency, lower cost, lighter footprint, and greater scalability than relays.
	Using \gls{ris} for {passive beamforming} has attracted significant interest in wireless communication \cite{Wu2019,Wu2020c,Yang2020,Zheng2021}, backscatter \cite{Jia2020,Liang2022}, sensing \cite{Liu2022a,Hua2023}, and power transfer literature \cite{Wu2021d,Feng2022,Zhao2022}, reporting a second-order array gain and fourth-order power scaling law (with proper waveform).
	On the other hand, \gls{ris} also enables {backscatter modulation} by dynamically switching between different patterns, as already investigated \cite{Karasik2020,Basar2020,Zhao2022a} and prototyped \cite{Tang2019b,Dai2020a}.
	Despite fruitful outcomes, one critical unanswered question is the {channel shaping} capability: \emph{To what extent can a passive \gls{ris} reshape the wireless channel?}

	The answer indeed depends on the hardware architecture and scattering model.
	In conventional (a.k.a. diagonal) \gls{ris}, each scattering element is tuned by a dedicated impedance and acts as an \emph{individual} phase shifter \cite{Wu2020}.
	The concept is generalized to \gls{bd}-\gls{ris} \cite{Shen2020a,Li2023b} which groups adjacent elements using passive components.
	This allows \emph{cooperative} scattering --- wave impinging on one element can propagate within the circuit and depart partially from any element in the same group.
	\gls{bd}-\gls{ris} can thus control both amplitude and phase of the reflected wave, generalizing the scattering matrix from diagonal with unit-magnitude entries to block diagonal with  unitary blocks.
	Its benefit has been recently shown in receive power maximization \cite{Nerini2023,Santamaria2023,Fang2023,Nerini2023a}, transmit power minimization \cite{Zhou2023}, and rate maximization \cite{Zhou2023,Nerini2023a,Li2023d,Bartoli2023,Li2023c}.
	Practical issues such as channel estimation \cite{Li2023e} and mutual coupling \cite{Li2023f} have also been investigated.
	Therefore, \gls{bd}-\gls{ris} is envisioned as the next generation channel shaper with stronger signal processing flexibility \cite{Li2023g}.

	% Attempts to characterize the channel shaping capability can be classified into \emph{singular value centric} and \emph{power centric} methods.
	% Channel shaping is different from passive beamforming as it seeks to modify the inherent properties of the channel itself, allowing one to decouple \gls{ris} and transceiver design.
	Channel shaping is different from passive beamforming as it seeks to modify the inherent properties of the channel itself.
	This allows one to decouple the \gls{ris}-transceiver design and explore the fundamental limits of channel manipulation.
	For example, diagonal \gls{ris} has been proved useful for improving channel power \cite{Ning2020}, degree of freedom \cite{Ozdogan2020a,Li2023h}, condition number \cite{Zheng2022,Huang2023}, and effective rank \cite{ElMossallamy2021,Meng2023} in \gls{mimo}.
	In contrast, \gls{bd}-\gls{ris} can provide a higher channel power but existing results are limited to \gls{siso}\footnote{In terms of channel shaping, single-stream \gls{mimo} with given precoder and combiner \cite{Nerini2023} is equivalent to \gls{siso}.}. \cite{Nerini2023} and \gls{miso} \cite{Santamaria2023}.
	While these studies offer promising glimpses into the channel shaping potential, a comprehensive understanding of the capabilities and limitations is desired, and a universal design framework is missing.
	This paper aims to answer the channel shaping question through theoretical analysis and numerical optimization.
	The contributions are summarized below.

	First, we quantify the capability of a \gls{bd}-\gls{ris} to reshape the \gls{mimo} point-to-point channel in terms of singular values.
	The \emph{Pareto frontiers} are characterized by optimizing the {weighted sum of singular values}, where the weights can be positive, zero, or negative.
	% This generalizes most singular value metrics and provides a powerful design framework.
	The resulting singular value region generalizes most relevant metrics and provides an intuitive channel shaping benchmark.
	We then discuss some analytical singular value bounds in rank-deficient and fully-connected scenarios, which help to demystify the gain from off-diagonal entries.
	This is the first paper to answer the channel shaping question and highlight the \gls{bd}-\gls{ris} gain from a Pareto perspective.

	Second, we propose a \gls{rcg} algorithm adapted from \cite{Abrudan2008,Abrudan2009} for smooth optimization problems of asymmetric \gls{bd}-\gls{ris} with arbitrary group size.
	Specifically, block-wise update is performed along the geodesics\footnote{A geodesic refers to the shortest path between two points in a Riemannian manifold.} of the Stiefel manifold, which are expressed compactly by the exponential map \cite{Edelman1998}.
	It features lower complexity and faster convergence than general manifold optimization \cite{Absil2009,Pan2022d}, and solves the Pareto singular value problem.
	This is the first paper to tailor an efficient optimization framework for asymmetric \gls{bd}-\gls{ris}.

	Third, we tackle \gls{bd}-\gls{ris} \gls{mimo} rate maximization with two solutions: a local-optimal approach through \gls{ao} and a low-complexity approach over channel shaping.
	The former updates active and passive beamformers by eigenmode transmission and \gls{rcg} algorithm, respectively.
	The latter suboptimally decouples both blocks, recasts the shaping problem as channel power maximization, and solves it iteratively in closed form.
	Interestingly, the gap in between vanishes as \gls{bd}-\gls{ris} evolves from diagonal (single-connected) to unitary (fully-connected).
	It suggests channel shaping offers a promising low-complexity solution for joint \gls{ris}-transceiver designs.

	% Third, we propose a closed-form iterative algorithm for power centric channel shaping problems.
	% The idea is to successively approximate the quadratic objective by a sequence of affines and solve the local problems by \gls{svd}.
	% Case studies are conducted for channel power maximization in \gls{pc} and leakage interference minimization in \gls{ic}.

	Fourth, extensive simulations reveal that the performance gain from \gls{bd}-\gls{ris} increases with group size and \gls{mimo} dimensions.
	In terms of channel power, fully-connected \gls{bd}-\gls{ris} boosts up to 62\% and 270\% over diagonal \gls{ris} in $1 \times 1$ and $4 \times 4$ \gls{mimo} under independent Rayleigh fading, respectively.
	The superiority stems from stronger \emph{subchannel rearrangement} and \emph{subspace alignment} capabilities empowered by in-group cooperation.
	It emphasizes the importance of using \gls{bd}-\gls{ris} in large-scale \gls{mimo} systems.

\end{section}

\begin{section}{\glsfmtlong{bd}-\glsfmtlong{ris} Model}
	Consider a \gls{bd}-\gls{ris} aided point-to-point \gls{mimo} system with $N_\mathrm{T}$, $N_\mathrm{S}$, $N_\mathrm{R}$ transmit, scatter, and receive antennas, respectively.
	This configuration is denoted as $N_\mathrm{T} \times N_\mathrm{S} \times N_\mathrm{R}$ in the following context.
	The \gls{bd}-\gls{ris} can be modeled as an $N_\mathrm{S}$-port network \cite{Ivrlac2010} that divides into $G$ individual groups, each containing $L \triangleq N_\mathrm{S} / G$ elements interconnected by real-time reconfigurable components \cite{Shen2020a}.
	To simplify the analysis, we assume a lossless asymmetric network\footnote{While symmetric \gls{bd}-\gls{ris} (involving capacitors and inductors) is often considered in the literature \cite{Shen2020a,Nerini2023,Santamaria2023,Fang2023,Nerini2023a,Zhou2023,Li2023d,Bartoli2023}, asymmetric \gls{bd}-\gls{ris} can be built over reconfigurable asymmetric passive components (e.g., ring hybrids and branch-line hybrids) \cite{Ahn2006}.} without mutual coupling between scatter elements, as previously considered in \cite{Li2023b,Li2023c,Bartoli2023}.
	The overall scattering matrix of the \gls{bd}-\gls{ris} is
	\begin{equation}
		\mathbf{\Theta} = \mathrm{diag}(\mathbf{\Theta}_1,\ldots,\mathbf{\Theta}_G),
	\end{equation}
	where $\mathbf{\Theta}_g \in \mathbb{U}^{L \times L}$ is a unitary matrix that describes the response of group $g \in \mathcal{G} \triangleq \{1, \ldots, G\}$.
	That is, $\mathbf{\Theta}_g = \mathbf{\Theta}_g^\mathsf{H}$ and $\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g = \mathbf{I}$.
	Note that diagonal (single-connected) and unitary (fully-connected) \gls{ris} can be regarded as its extreme cases with group size $L=1$ and $L=N_\mathrm{S}$, respectively.
	Some potential physical architectures of \gls{bd}-\gls{ris} are illustrated in \cite[Fig. 3]{Shen2020a}, \cite[Fig. 5]{Li2023c}, and \cite[Fig. 2]{Nerini2023a}, where the radiation pattern and circuit topology have impacts on the scattering matrix.

	Let $\mathbf{H}_\mathrm{D} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{B} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{S}}$, $\mathbf{H}_\mathrm{F} \in \mathbb{C}^{N_\mathrm{S} \times N_\mathrm{T}}$ denote the direct (transmitter-receiver), backward (\gls{ris}-receiver), and forward (transmitter-\gls{ris}) channels, respectively.
	The equivalent channel is a function of the scattering matrix
	\begin{equation}
		\mathbf{H} = \mathbf{H}_\mathrm{D} + \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F} = \mathbf{H}_\mathrm{D} + \sum_g \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g},
		\label{eq:channel_equivalent}
	\end{equation}
	where $\mathbf{H}_{\mathrm{B},g} \in \mathbb{C}^{N_\mathrm{R} \times L}$ and $\mathbf{H}_{\mathrm{F},g} \in \mathbb{C}^{L \times N_\mathrm{T}}$ are the backward and forward subchannels for \gls{ris} group $g$, corresponding to the $(g-1)L$ to $gL$ columns of $\mathbf{H}_\mathrm{B}$ and rows of $\mathbf{H}_\mathrm{F}$, respectively.
	% That is, $\mathbf{H}_{\mathrm{B},g}$ (resp. $\mathbf{H}_{\mathrm{F},g}$) corresponds to the $(g-1)L$ to $gL$ columns (resp. rows) of $\mathbf{H}_\mathrm{B}$ (resp. $\mathbf{H}_\mathrm{F}$).
	% , corresponding to the $(g-1)L$ to $gL$ columns of $\mathbf{H}_\mathrm{B}$ and rows of $\mathbf{H}_\mathrm{F}$.
	% $\mathbf{H}_{\mathrm{B},g}$ (resp. $\mathbf{H}_{\mathrm{F},g}$) corresponds to the $(g-1)L$ to $gL$ columns of $\mathbf{H}_\mathrm{B}$
	Since unitary matrices constitute an algebraic group with respect to multiplication, the scattering matrix of group $g$ can be decomposed as
	\begin{equation}
		\mathbf{\Theta}_g = \mathbf{L}_g \mathbf{R}_g^\mathsf{H},
	\end{equation}
	where $\mathbf{L}_g, \mathbf{R}_g \in \mathbb{U}^{L \times L}$ are two unitary factor matrices.
	Let $\mathbf{H}_{\mathrm{B},g} = \mathbf{U}_{\mathrm{B},g} \mathbf{\Sigma}_{\mathrm{B},g} \mathbf{V}_{\mathrm{B},g}^\mathsf{H}$ and $\mathbf{H}_{\mathrm{F},g} = \mathbf{U}_{\mathrm{F},g} \mathbf{\Sigma}_{\mathrm{F},g} \mathbf{V}_{\mathrm{F},g}^\mathsf{H}$ be the compact \gls{svd} of the backward and forward channels, respectively.
	The equivalent channel can thus be rewritten as
	\begin{equation}
		\mathbf{H} = \overbrace{\mathbf{H}_\mathrm{D} + \sum_g \mathbf{U}_{\mathrm{B},g} \mathbf{\Sigma}_{\mathrm{B},g} \underbrace{\mathbf{V}_{\mathrm{B},g}^\mathsf{H} \mathbf{L}_g \mathbf{R}_g^\mathsf{H} \mathbf{U}_{\mathrm{F},g}}_\text{backward-forward} \mathbf{\Sigma}_{\mathrm{F},g} \mathbf{V}_{\mathrm{F},g}^\mathsf{H}}^\text{direct-indirect}.
		\label{eq:channel_equivalent_svd}
	\end{equation}

	\begin{remark}
		By analyzing \eqref{eq:channel_equivalent_svd}, we conclude that channel shaping through \gls{bd}-\gls{ris} creates two critical opportunities:
		\begin{itemize}
			\item \textbf{Subspace alignment:} each group can align the singular matrices of the backward-forward (intra-group, multiplicative) and the direct-indirect (inter-group, additive) channels. In \gls{siso}, subspace alignment boils down to phase matching and the optimal scattering matrix of group $g$ is
			\begin{equation}
				\mathbf{\Theta}_g^\star = \exp \bigl(\jmath \mathrm{arg}(h_\mathrm{D})\bigr) \mathbf{V}_{\mathrm{B},g} \mathbf{U}_{\mathrm{F},g}^\mathsf{H},
			\end{equation}
			where $\mathbf{V}_{\mathrm{B},g} = \bigl[\mathbf{h}_{\mathrm{B},g}/\lVert \mathbf{h}_{\mathrm{B},g} \rVert, \mathbf{N}_{\mathrm{B},g}\bigr]$, $\mathbf{U}_{\mathrm{F},g} = \bigl[\mathbf{h}_{\mathrm{F},g}/\lVert \mathbf{h}_{\mathrm{F},g} \rVert, \mathbf{N}_{\mathrm{F},g}\bigr]$, and $\mathbf{N}_{\mathrm{B},g}, \mathbf{N}_{\mathrm{F},g} \in \mathbb{C}^{L \times (L-1)}$ are the orthonormal bases of the null spaces of $\mathbf{h}_{\mathrm{B},g}$ and $\mathbf{h}_{\mathrm{F},g}$, respectively.
			Diagonal \gls{ris} is thus sufficient for perfect \gls{siso} phase matching (with empty null spaces), but its disadvantage in subspace alignment scales with \gls{mimo} dimensions, since each element can only apply a common phase shift to the ``pinhole'' indirect channel of size $N_\mathrm{R} \times N_\mathrm{T}$ passing through itself.
			Even if the \gls{bd}-\gls{ris} is fully-connected, there still exists a tradeoff between the direct-indirect and backward-forward alignments.
			% It suggests that diagonal \gls{ris} applying a phase shift to each indirect channel is sufficient for phase matching in \gls{siso}. but its disadvantage in subspace alignment scales with \gls{mimo} dimensions.
			% It suggests that diagonal \gls{ris} applying a phase shift to each indirect channel is sufficient for phase matching in \gls{siso}, but its disadvantage in subspace alignment scales with \gls{mimo} dimensions.
			% It is obvious that diagonal \gls{ris} is sufficient for perfect phase matching. However,
			% It is obvious that diagonal \gls{ris} is sufficient for perfect phase matching. However, in \gls{mimo} each element can only apply a common phase shift to the associated rank-1 $N_\mathrm{R} \times N_\mathrm{T}$ indirect channel.
			% Therefore, perfect subspace alignment of indirect channels through different elements is generally impossible.
			% It means the disadvantage of diagonal \gls{ris} in subspace alignment and subchannel rearrangement scales with \gls{mimo} dimensions.
			% is sufficient for perfect alignment
			% and is sufficient for perfect subspace alignment.
			% In \gls{mimo}, perfect subspace alignment is generally impossible due to the lack of direct control over the indirect channels.aa
			% backward-forward (intra-group, multiplicative) and the direct-indirect (inter-group, additive) alignments of the singular matrices.
			\item \textbf{Subchannel rearrangement:} this unique property of \gls{bd}-\gls{ris} allows each group to rearrange and combine the backward and forward subchannels by strength. In \gls{siso}, diagonal \gls{ris} provides a maximum indirect channel amplitude of $\sum_{n=1}^{N_\mathrm{S}} \lvert h_{\mathrm{B},n} \rvert \lvert h_{\mathrm{F},n} \rvert$, while \gls{bd}-\gls{ris} can generalize it to $\sum_{g=1}^{G} \sum_{l=1}^{L} \lvert h_{\mathrm{B},\pi_{\mathrm{B},g}(l)} \rvert \lvert h_{\mathrm{F},\pi_{\mathrm{F},g}(l)} \rvert$, where $\pi_{\mathrm{B},g}$ and $\pi_{\mathrm{F},g}$ are permutations of $\mathcal{L} \triangleq \{1, \ldots, L\}$. By rearrangement inequality, the maximum is attained by pairing the $l$-th strongest backward and forward subchannels in each group.
			The advantage of \gls{bd}-\gls{ris} in subchannel rearrangement scales with \gls{mimo} dimensions, since the number of subchannels associated with each group grows with $N_\mathrm{R} \times N_\mathrm{T}$.
			% effect is more pronounced in \gls{mimo} since the number of subchannels associated with each group grows quadratically with the dimensions.
			% The advantage of \gls{bd}-\gls{ris} in subchannel rearrangement scales with \gls{mimo} dimensions. since the number of subchannels associated with each group grows quadratically with the dimensions.
			% It suggests that the advantage of \gls{bd}-\gls{ris} in subchannel rearrangement scales with \gls{mimo} dimensions.
			% \item \textbf{Subchannel rearrangement:} this unique property of \gls{bd}-\gls{ris} allows each group to rearrange and combine the backward and forward subchannels by strength. In \gls{siso}, the maximum indirect channel amplitude by diagonal \gls{ris} is $\sum_{n=1}^{N_\mathrm{S}} \lvert h_{\mathrm{B},n} \rvert \lvert h_{\mathrm{F},n} \rvert$ while \gls{bd}-\gls{ris} can generalize it to $\sum_{g=1}^{G} \sum_{l=1}^{L} \lvert h_{\mathrm{B},\pi_{\mathrm{B},g}(l)} \rvert \lvert h_{\mathrm{F},\pi_{\mathrm{F},g}(l)} \rvert$, where $\pi_{\mathrm{B},g}$ and $\pi_{\mathrm{F},g}$ are independent permutations of $\mathcal{L} \triangleq \{1, \ldots, L\}$. By rearrangement inequality, the maximum is attained by pairing the $l$-th strongest backward and forward subchannels in each group.
			% \begin{equation}
			% 	\item \
			% \end{equation}
			% and $\sum_{g=1}^{G} \sum_{l=1}^{L} \lvert h_{\mathrm{B},\pi_{\mathrm{B},g}(l)} \rvert \lvert h_{\mathrm{F},\pi_{\mathrm{F},g}(l)} \rvert$ by \gls{bd}-\gls{ris}, where $\pi_{\mathrm{B},g}$ and $\pi_{\mathrm{F},g}$ are permutations of $\mathcal{L} \triangleq \{1, \ldots, L\}$. By rearrangement inequality,
			% \item \textbf{Subchannel rearrangement:} this unique property of \gls{bd}-\gls{ris} allows each group to rearrange and combine the backward and forward subchannels by strength. In \gls{siso}, the optimal indirect channel amplitude is $\sum_{n=1}^{N_\mathrm{S}} \lvert h_{\mathrm{B},n} \rvert \lvert h_{\mathrm{F},n} \rvert$ by diagonal \gls{ris} and $\sum_{n=1}^{N_\mathrm{S}} \lvert h_{\mathrm{B},\pi_\mathrm{B}(n)} \rvert \lvert h_{\mathrm{F},\pi_\mathrm{F}(n)} \rvert$ by \gls{bd}-\gls{ris}, where $\pi(\cdot)$ is a permutation of $\{1, \ldots, N_\mathrm{S}\}$.
			% this is unique to \gls{bd}-\gls{ris} --- each group can rearrange and combine the backward and forward subchannels by strength. In \gls{siso}, the optimal indirect channel amplitude is relaxed from $\sum_n \lvert h_{\mathrm{B},n} \rvert \lvert h_{\mathrm{F},n} \rvert$ to $\sum_n \lvert h_{\mathrm{B},\pi(n)} \rvert \lvert h_{\mathrm{F},\pi(n)} \rvert$ where $\pi(\cdot)$ is a permutation of $\{1, \ldots, L\}$.
			% from strongest to weakest, which attains the maximum in rearrangement inequality.
		\end{itemize}
	\end{remark}

	% TODO: add illustrations for both

	% In \gls{siso}, the equivalent channel can be further simplified as
	% \begin{equation}
	% 	h = h_\mathrm{D} + \sum_g \mathbf{h}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{h}_{\mathrm{F},g},
	% \end{equation}
	% subspace alignment boils down to phase matching

	% For the \gls{siso} case in Fig. \subref*{fg:power_bond_txrx1_nd}, the maximum channel power is approximately \num{4e-6} by diagonal \gls{ris} and \num{6.5e-6} by fully-connected \gls{bd}-\gls{ris}, corresponding to a \qty{62.5}{\percent} gain.
	% This aligns with the asymptotic \gls{bd}-\gls{ris} scaling law derived for \gls{siso} in \cite{Shen2020a}.
	% Interestingly, the gain surges to \qty{270}{\percent} in 4T4R \gls{mimo} as shown in Fig. \subref*{fg:power_bond_txrx4_nd}.
	% This is because subspace alignment boils down to phase matching in \gls{siso} such that both triangular and Cauchy-Schwarz inequalities in \cite[(50)]{Shen2020a} can be simultaneously tight regardless of the group size.
	% That is, diagonal \gls{ris} is sufficient for subspace alignment in \gls{siso} while the \qty{62.5}{\percent} gain from \gls{bd}-\gls{ris} comes purely from subchannel rearrangement (i.e., pairing the forward and backward channels from strongest to weakest).
	% Now consider a diagonal \gls{ris} in \gls{mimo}.
	% Each element can only apply a common phase shift to the associated rank-1 $N_\mathrm{R} \times N_\mathrm{T}$ indirect channel.
	% Therefore, perfect subspace alignment of indirect channels through different elements is generally impossible.
	% It means the disadvantage of diagonal \gls{ris} in subspace alignment and subchannel rearrangement scales with \gls{mimo} dimensions.
	% We thus conclude that the power gain of \gls{bd}-\gls{ris} scales with group size and \gls{mimo} dimensions.

	% \begin{equation}
	% 	\mathbf{H} = \underbrace{\mathbf{H}_\mathrm{D} + \underbrace{\sum_g \mathbf{U}_{\mathrm{B},g} \mathbf{\Sigma}_{\mathrm{B},g} \underbrace{\mathbf{V}_{\mathrm{B},g}^\mathsf{H} \mathbf{L}_g \mathbf{R}_g^\mathsf{H} \mathbf{U}_{\mathrm{F},g}}_\text{backward-forward} \mathbf{\Sigma}_{\mathrm{F},g} \mathbf{V}_{\mathrm{F},g}^\mathsf{H}}_\text{group-wise}}_\text{direct-indirect},
	% \end{equation}
	% which involves three subspace alignment problems:
	% which involves three subspace alignment problems
	% \begin{equation}
	% 	\mathbf{H} = \mathbf{H}_\mathrm{D} + \sum_g \mathbf{U}_{\mathrm{B},g} \mathbf{\Sigma}_{\mathrm{B},g} \underbrace{\mathbf{V}_{\mathrm{B},g}^\mathsf{H} \mathbf{L}_g \mathbf{R}_g^\mathsf{H} \mathbf{U}_{\mathrm{F},g}}_{\substack{\text{backward-forward}\\\text{additive alignment}}} \mathbf{\Sigma}_{\mathrm{F},g} \mathbf{V}_{\mathrm{F},g}^\mathsf{H}.
	% \end{equation}
	% \begin{equation}
	% 	\mathbf{H} = \mathbf{H}_\mathrm{D} + \sum_g \mathbf{U}_{\mathrm{B},g} \mathbf{\Sigma}_{\mathrm{B},g} \underbrace{\mathbf{V}_{\mathrm{B},g}^\mathsf{H} \mathbf{L}_g}_{\substack{\text{backward}\\\text{align}}} \underbrace{\mathbf{R}_g^\mathsf{H} \mathbf{U}_{\mathrm{F},g}}_{\substack{\text{forward}\\\text{align}}} \mathbf{\Sigma}_{\mathrm{F},g} \mathbf{V}_{\mathrm{F},g}^\mathsf{H}.
	% \end{equation}
	% Then the equivalent channel can be rewritten as
	% Since the field of (block) unitary matrices formulate a field that is closed under multiplication
	% Since block unitary matrices satisfy the axioms of a field under matrix multiplication
	% \begin{remark}
	% 	% To understand the \gls{b} gain from in-group connections
	% \end{remark}

	% \begin{remark}
	% 	From \eqref{eq:scattering_fc} and \eqref{eq:channel_equivalent_fc} in the proof of Proposition \ref{pp:fully_connected}, we notice that \eqref{iq:sv_bound_fc}--\eqref{iq:sv_bound_fc_power} are simultaneously tight when
	% 	% Interestingly, \eqref{iq:sv_bound_fc}--\eqref{iq:sv_bound_fc_power} are simultaneously tight when
	% 	\begin{equation}
	% 		\mathbf{\Theta} = \mathbf{V}_\mathrm{B} \mathbf{U}_\mathrm{F}^\mathsf{H}.
	% 		\label{eq:scattering_fc_tight}
	% 	\end{equation}
	% 	An interpretation is that the off-diagonal entries can enhance the capabilities of
	% 	\begin{itemize}
	% 		\item subspace alignment: $\mathbf{V}_\mathrm{B}$ and $\mathbf{U}_\mathrm{F}^\mathsf{H}$ in \eqref{eq:scattering_fc} fully align the subspaces of $\mathbf{H}_\mathrm{B}$ and $\mathbf{H}_\mathrm{F}$ by rotation;
	% 		\item subchannel rearrangement: $\mathbf{X} = \mathbf{I}$ in \eqref{eq:channel_equivalent_fc} pairs the subchannels of $\mathbf{H}_\mathrm{B}$ and $\mathbf{H}_\mathrm{F}$ from strongest to weakest, which attains the maximum in rearrangement inequality.
	% 	\end{itemize}
	% \end{remark}

	% \begin{remark}
	% 	\gls{bd}-\gls{ris} reduces to diagonal \gls{ris} and unitary \gls{ris} with group size $L$ 1 and $N_\mathrm{S}$, respectively.
	% \end{remark}
	% \begin{remark}
	% 	Individual forward and backward \gls{csi} are required for \gls{bd}-\gls{ris} designs.
	% 	This is different from diagonal \gls{ris} where estimating their product is usually sufficient.
	% 	% Later we will show the potential benefits from the \gls{csi} overhead.
	% \end{remark}
\end{section}

\begin{section}{Channel Singular Values Redistribution}
	\begin{subsection}{A Toy Example}\label{sc:toy_example}
		We first illustrate the channel shaping capabilities of different \gls{ris} by a toy example.
		% We first use a toy example to illustrate that \gls{bd}-\gls{ris} can provide a wider dynamic range of channel singular values.
		Consider a $2 \times 2 \times 2$ setup where the direct link is blocked.
		The diagonal \gls{ris} is modeled by $\mathbf{\Theta}_\mathrm{D} = \mathrm{diag}(e^{\jmath \theta_1}, e^{\jmath \theta_2})$ while the unitary \gls{bd}-\gls{ris} has 4 independent angular parameters
		\begin{equation}
			\mathbf{\Theta}_\mathrm{U} = e^{\jmath \phi} \begin{bmatrix}
				e^{\jmath \alpha} \cos \psi  & e^{\jmath \beta} \sin \psi   \\
				-e^{-\jmath \beta} \sin \psi & e^{-\jmath \alpha} \cos \psi
			\end{bmatrix}.
		\end{equation}
		In particular, $\phi$ has no impact on the singular value because $\mathrm{sv}(e^{\jmath \phi} \mathbf{A}) = \mathrm{sv}(\mathbf{A})$.
		We also enforce symmetry by $\beta = \pi / 2$ such that both architectures have the same number of angular parameters.
		\begin{figure}
			\centering
			\includegraphics[width=\columnwidth]{assets/chapter_5/singular_trend.eps}
			\caption{$2 \times 2 \times 2$ (no direct) channel singular value shaping by diagonal and symmetry unitary \gls{ris}.}
			\label{fg:singular_trend}
		\end{figure}
		Fig.~\ref{fg:singular_trend} shows the channel singular values achieved by an exhaustive grid search over $(\theta_1, \theta_2)$ for diagonal \gls{ris} and $(\alpha, \psi)$ for symmetric unitary \gls{ris}.
		It is observed that both singular values can be manipulated up to $9\%$ using diagonal \gls{ris} and $42\%$ using symmetric \gls{bd}-\gls{ris}, despite both architectures have the same number of scattering elements and design parameters.
		A larger performance gap is expected when asymmetric \gls{bd}-\gls{ris} is available.
		This example shows \gls{bd}-\gls{ris} can provide a wider dynamic range of channel singular values and motivates further studies on channel shaping.
	\end{subsection}

	\begin{subsection}{Pareto Frontier Characterization}
		We then characterize the Pareto frontier of channel singular values by maximizing their weighted sum
		\begin{maxi!}
			{\scriptstyle{\mathbf{\Theta}}}{\sum_n \rho_n \sigma_n(\mathbf{H})}{\label{op:pareto}}{\label{ob:pareto}}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I},}{\quad \forall g,}{\label{co:pareto_unitary}}
		\end{maxi!}
		where $n \in \{1,\ldots,N \triangleq \min(N_\mathrm{T}, N_\mathrm{R})\}$ and $\rho_n$ is the weight of the $n$-th singular value that can be positive, zero, or negative.
		Varying $\{\rho_n\}$ unveils the entire achievable singular value region.
		Thus, the Pareto frontier problem~\eqref{op:pareto} generalizes most relevant metrics and provides a powerful shaping framework.
		The objective \eqref{ob:pareto} is smooth in $\mathbf{\Theta}$ and the feasible domain \eqref{co:pareto_unitary} for group $g$ corresponds to the Stiefel manifold.
		Next, we zoom out to general smooth maximization problems of asymmetric \gls{bd}-\gls{ris}.

		Inspired by \cite{Abrudan2008,Abrudan2009}, we propose a block-wise \gls{rcg} algorithm along the geodesics on the Lie group of unitary matrices $\mathbb{U}^{L \times L}$.
		It leverages the fact that unitary matrices are closed under multiplication.
		At iteration $r$, the gradient is computed in the Euclidean space and translated to the Riemannian manifold \cite{Absil2009}
		\begin{gather}
			\nabla_{\mathrm{E},g}^{(r)} = \frac{\partial f(\mathbf{\Theta}_g^{(r)})}{\partial \mathbf{\Theta}_g^*},\label{eq:gradient_euclidean}\\
			\nabla_{\mathrm{R},g}^{(r)} = \nabla_{\mathrm{E},g}^{(r)} {\mathbf{\Theta}_g^{(r)}}^\mathsf{H} - \mathbf{\Theta}_g^{(r)} {\nabla_{\mathrm{E},g}^{(r)}}^\mathsf{H}.\label{eq:gradient_riemannian} % $\mathcal{O}(N_\mathrm{S}^3)$
		\end{gather}
		The Polak-Ribierre parameter \cite{Polak1969} is approximated as \cite{Abrudan2009}
		\begin{equation}
			\gamma_g^{(r)} = \frac{\mathrm{tr}\bigl((\nabla_{\mathrm{R},g}^{(r)} - \nabla_{\mathrm{R},g}^{(r-1)}) {\nabla_{\mathrm{R},g}^{(r)}}^\mathsf{H}\bigr)}{\mathrm{tr}\bigl(\nabla_{\mathrm{R},g}^{(r-1)} {\nabla_{\mathrm{R},g}^{(r-1)}}^\mathsf{H}\bigr)}, % $\mathcal{O}(2 N_\mathrm{S}^3 + N_\mathrm{S}^2 + 2 N_\mathrm{S})$
			\label{eq:parameter_cg}
		\end{equation}
		and the conjugate direction is
		\begin{equation}
			\mathbf{D}_g^{(r)} = \nabla_{\mathrm{R},g}^{(r)} + \gamma_g^{(r)} \mathbf{D}_g^{(r-1)}. % $\mathcal{O}(N_\mathrm{S}^2)$
			\label{eq:direction_cg}
		\end{equation}
		In the Stiefel manifold, the geodesic emanating from $\mathbf{\Theta}_g^{(r)}$ with velocity $\mathbf{D}_g^{(r)}$ and step size $\mu$ is described compactly by the exponential map \cite{Edelman1998}
		\begin{equation}
			\mathbf{G}_g^{(r)}(\mu) = \exp(\mu \mathbf{D}_g^{(r)}) \mathbf{\Theta}_g^{(r)}. % $\mathcal{O}(N_\mathrm{S}^3)$
			\label{eq:geodesic}
		\end{equation}
		An appropriate $\mu^\star$ can be obtained by the Armijo rule \cite{Armijo1966}.\footnote{To double the step size, one only need to square the rotation matrix instead of recomputing the matrix exponential, i.e., $\exp(2 \mu \mathbf{D}_g^{(r)}) = \exp^2(\mu \mathbf{D}_g^{(r)})$.}
		Finally, the scattering matrix is updated along the geodesic as
		\begin{equation}
			\mathbf{\Theta}_g^{(r+1)} = \mathbf{G}_g^{(r)}(\mu^\star).
			\label{eq:update_geodesic}
		\end{equation}
		Algorithm~\ref{ag:rcg} summarizes the proposed block-wise geodesic \gls{rcg} method for smooth maximization problems of asymmetric \gls{bd}-\gls{ris}.
		Convergence to stationary points is guaranteed.

		\begin{remark}
			Compared with universal manifold optimization \cite{Absil2009,Pan2022d}, Algorithm~\ref{ag:rcg} inherits a trifold benefit from \cite{Abrudan2008,Abrudan2009}:
			\begin{enumerate}
				\item No retraction thanks to rotational update \eqref{eq:geodesic}, \eqref{eq:update_geodesic};
				\item Lower computational complexity per iteration;
				\item Faster convergence thanks to proper parameter space.
			\end{enumerate}
		\end{remark}

		% \setalgorithmcaptionfont{\small}
		\begin{algorithm}[!t]
			% \small
			\caption{Block-wise geodesic \gls{rcg} for asymmetric \gls{bd}-\gls{ris}}
			\label{ag:rcg}
			\begin{algorithmic}[1]
				\Require $f(\mathbf{\Theta})$, $G$
				\Ensure $\mathbf{\Theta}^\star$
				\Initialize {$r \gets 0$, $\mathbf{\Theta}^{(0)}$}
				\Repeat
					\For {$g \gets 1$ to $G$}
						\State $\nabla_{\mathrm{E},g}^{(r)} \gets$ \eqref{eq:gradient_euclidean} \label{ln:gradient_euclidean}
						\State $\nabla_{\mathrm{R},g}^{(r)} \gets$ \eqref{eq:gradient_riemannian}
						\State $\gamma_g^{(r)} \gets$ \eqref{eq:parameter_cg}
						\State $\mathbf{D}_g^{(r)} \gets$ \eqref{eq:direction_cg}
						\If {$\Re\bigl\{\mathrm{tr}({\mathbf{D}_g^{(r)}}^\mathsf{H} \nabla_{\mathrm{R},g}^{(r)})\bigr\} < 0$} \Comment{not an ascent direction}
							\State $\mathbf{D}_g^{(r)} \gets \nabla_{\mathrm{R},g}^{(r)}$
						\EndIf
						\State $\mu \gets 1$
						\State $\mathbf{G}_g^{(r)}(\mu) \gets$ \eqref{eq:geodesic}
						\While {$f\bigl(\mathbf{G}_g^{(r)}(2\mu)\bigr) - f(\mathbf{\Theta}_g^{(r)}) \ge \mu \cdot \mathrm{tr}(\mathbf{D}_g^{(r)} {\mathbf{D}_g^{(r)}}^\mathsf{H}) / 2$}
							\State $\mu \gets 2 \mu$
						\EndWhile
						\While {$f\bigl(\mathbf{G}_g^{(r)}(\mu)\bigr) - f(\mathbf{\Theta}_g^{(r)}) < \mu / 2 \cdot \mathrm{tr}(\mathbf{D}_g^{(r)} {\mathbf{D}_g^{(r)}}^\mathsf{H}) / 2$}
							\State $\mu \gets \mu / 2$
						\EndWhile
						\State $\mathbf{\Theta}_g^{(r+1)} \gets$ \eqref{eq:update_geodesic}
					\EndFor
					\State $r \gets r+1$
				\Until $\lvert f(\mathbf{\Theta}^{(r)}) - f(\mathbf{\Theta}^{(r-1)}) \rvert / f(\mathbf{\Theta}^{(r-1)}) \le \epsilon$
			\end{algorithmic}
		\end{algorithm}

		\begin{lemma}\label{lm:pareto_gradient}
			The Euclidean gradient of \eqref{ob:pareto} w.r.t. \gls{bd}-\gls{ris} group $g$ is
			\begin{equation}
				\frac{\partial \sum_n \rho_n \sigma_n(\mathbf{H})}{\partial \mathbf{\Theta}_g^*} = \mathbf{H}_{\mathrm{B},g}^\mathsf{H} \mathbf{U} \mathrm{diag}(\rho_1,\ldots,\rho_N) \mathbf{V}^\mathsf{H} \mathbf{H}_{\mathrm{F},g}^\mathsf{H},
				\label{eq:pareto_gradient}
			\end{equation}
			where $\mathbf{U}$ and $\mathbf{V}$ are the left and right singular matrices of $\mathbf{H}$, respectively.
		\end{lemma}
		\begin{proof}
			Please refer to Appendix~\ref{ap:pareto_gradient}.
		\end{proof}

		%  The Euclidean gradient of \gls{bd}-\gls{ris} group $g$ is computed as
		% where $\mathbf{U} \in \mathbb{C}^{N_\mathrm{R} \times N}$ and $\mathbf{V} \in \mathbb{C}^{N_\mathrm{T} \times N}$ are the left and right compact singular matrices of $\mathbf{H}$, respectively.
		Algorithm~\ref{ag:rcg} can thus be invoked for the Pareto singular value problem \eqref{op:pareto} where line \ref{ln:gradient_euclidean} uses \eqref{eq:pareto_gradient} explicitly.
		% \eqref{eq:pareto_gradient} is used in line \ref{ln:gradient_euclidean}.
	\end{subsection}

	\begin{subsection}{Some Analytical Bounds}\label{sc:bounds}
		We then discuss some analytical bounds related to channel singular values.
		\begin{proposition}[degree of freedom]\label{pp:dof}
			In point-to-point \gls{mimo}, \gls{bd}-\gls{ris} cannot achieve a higher \gls{dof} than diagonal \gls{ris}.
		\end{proposition}
		\begin{proof}
			Please refer to Appendix~\ref{ap:dof}.
		\end{proof}

		\begin{proposition}[rank-deficient channel]\label{pp:rank_deficient}
			% If the forward/backward channel is rank-$k$, then a \gls{bd}-\gls{ris} can at most enlarge the $n$-th ($n > k$) channel singular value to the $n-k$-th singular value of $\mathbf{T}$, or suppress the $n$-th ($n < N - k + 1$) channel singular value to the $n$-th singular value of $\mathbf{T}$.
			% If the forward/backward channel is rank-$k$, then a sufficiently large \gls{bd}-\gls{ris} can at most enlarge (resp. suppress) the first (resp. last) $k$ equivalent channel singular values without bounds.
			If the forward or backward channel is rank-$k$, then regardless of the \gls{ris} size and architecture, the $n$-th singular value of the equivalent channel is bounded by
			\begin{align}
				\sigma_n(\mathbf{H}) & \le \sigma_{n-k}(\mathbf{T}), &  & \text{if } n > k, \label{iq:sv_bound_enlarge}          \\
				\sigma_n(\mathbf{H}) & \ge \sigma_n(\mathbf{T}),     &  & \text{if } n < N - k + 1, \label{iq:sv_bound_suppress}
			\end{align}
			where
			% \begin{align}
			% 	\mathbf{T} \mathbf{T}^\mathsf{H} = \mathbf{H}_\mathrm{D} (\mathbf{I} - \mathbf{V}_\mathrm{F} \mathbf{V}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}^\mathsf{H}, & \text{if } \mathrm{rank}(\mathbf{H}_\mathrm{F}) = k, \\
			% 	\mathbf{T}^\mathsf{H} \mathbf{T} = \mathbf{H}_\mathrm{D}^\mathsf{H} (\mathbf{I} - \mathbf{U}_\mathrm{F} \mathbf{U}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}, & \text{if } \mathrm{rank}(\mathbf{H}_\mathrm{B}) = k.
			% \end{align}
			\begin{equation}
				\mathbf{T} \mathbf{T}^\mathsf{H} =
				\begin{cases}
					\mathbf{H}_\mathrm{D} (\mathbf{I} - \mathbf{V}_\mathrm{F} \mathbf{V}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}^\mathsf{H}, & \text{if } \mathrm{rank}(\mathbf{H}_\mathrm{F}) = k, \\
					\mathbf{H}_\mathrm{D}^\mathsf{H} (\mathbf{I} - \mathbf{U}_\mathrm{B} \mathbf{U}_\mathrm{B}^\mathsf{H}) \mathbf{H}_\mathrm{D}, & \text{if } \mathrm{rank}(\mathbf{H}_\mathrm{B}) = k,
				\end{cases}
				\label{eq:auxiliary_matrix}
			\end{equation}
			and $\mathbf{V}_\mathrm{F}$ and $\mathbf{U}_\mathrm{B}$ are the right and left compact singular matrices of $\mathbf{H}_\mathrm{F}$ and $\mathbf{H}_\mathrm{B}$, respectively.
			% where $\mathbf{T} \mathbf{T}^\mathsf{H} = \mathbf{H}_\mathrm{D} (\mathbf{I} - \mathbf{V}_\mathrm{F} \mathbf{V}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}$ if the forward channel is rank-$k$ and $\mathbf{T}^\mathsf{H} \mathbf{T} = \mathbf{H}_\mathrm{D} (\mathbf{I} - \mathbf{U}_\mathrm{F} \mathbf{U}_\mathrm{F}^\mathsf{H}) \mathbf{H}_\mathrm{D}$ if the backward channel is rank-$k$.
		\end{proposition}
		\begin{proof}
			Please refer to Appendix~\ref{ap:rank_deficient}.
		\end{proof}

		% \begin{corollary}[degree of freedom]
		% 	In point-to-point \gls{mimo} with or without direct link, \gls{bd}-\gls{ris} cannot provide a higher \gls{dof} than diagonal \gls{ris}.
		% \end{corollary}

		\begin{corollary}[extreme singular values\label{co:extreme}]
			With a sufficiently large \gls{ris}, the first $k$ channel singular values are unbounded above while the last $k$ channel singular values can be suppressed to zero.
			% For a rank-$k$ forward or backward channel with a sufficiently large \gls{ris}, the first $k$ singular values of the equivalent channel are unbounded above while the last $k$ singular values can be suppressed to zero.
			% If the forward or backward channel is rank-$k$ and the \gls{ris} is sufficiently large, then the first $k$ singular values of the equivalent channel are unbounded above while the last $k$ singular values can be suppressed to zero.
			% If the forward/backward channel is rank-$k$, then the first (resp. last) $k$ singular values of the equivalent channel are unbounded above (resp. below).
			% a sufficiently large \gls{ris} can at most enlarge (resp. suppress) the first (resp. last) $k$ singular values of the equivalent channel without bounds.
		\end{corollary}

		\begin{corollary}[\gls{los} channel\footnote{A similar result has been derived for diagonal \gls{ris} in \cite{Semmler2023}.}\label{co:los}]
			If the forward or backward channel is \gls{los}, then a \gls{ris} can at most enlarge (resp. suppress) the $n$-th ($n \ge 2$) channel singular value to the $(n-1)$-th (resp. $n$-th) singular value of $\mathbf{T}$, that is,
			\begin{equation}
				\sigma_1(\mathbf{H}) \ge \sigma_1(\mathbf{T}) \ge {\sigma_2(\mathbf{H})} \ge \ldots \ge \sigma_{N-1}(\mathbf{T}) \ge {\sigma_N(\mathbf{H})} \ge \sigma_N(\mathbf{T}).
				\label{iq:sv_bound_los}
			\end{equation}
		\end{corollary}

		In Section~\ref{sc:simulation}, we will show by simulation that a finite-size \gls{bd}-\gls{ris} can approach those bounds better than diagonal \gls{ris}.

		\begin{proposition}[fully-connected \gls{ris} without direct link]\label{pp:fully_connected}
			% If the \gls{bd}-\gls{ris} is fully-connected and the direct link is absent, then the singular value bounds on $\mathbf{H}$ are equivalent to the singular value bounds on $\mathbf{BF}$, where $\mathbf{B}$ and $\mathbf{F}$ are arbitrary matrices with the same singular values as $\mathbf{H}_\mathrm{B}$ and $\mathbf{H}_\mathrm{F}$, respectively,
			If the \gls{bd}-\gls{ris} is fully-connected and the direct link is absent, then the channel singular values can be manipulated up to
			% If the \gls{bd}-\gls{ris} is fully-connected and the direct link is absent, then the bounds on the channel singular values are equivalent to the singular values of the equivalent channel, that is,
			% only bounds that apply here are those that apply to the singular values of
			\begin{equation}
				\mathrm{sv}(\mathbf{H}) = \mathrm{sv}(\mathbf{BF}),
			\end{equation}
			where $\mathbf{B}$ and $\mathbf{F}$ are arbitrary matrices with the same singular values as $\mathbf{H}_\mathrm{B}$ and $\mathbf{H}_\mathrm{F}$, respectively,
			% Our focus thus shifts to how the singular values of matrix product are bounded by the singular values of its individual factors.
		\end{proposition}

		\begin{proof}
			Please refer to Appendix~\ref{ap:fully_connected}.
		\end{proof}

		The problem now becomes how the singular values of matrix product are bounded by the singular values of its individual factors.
		Let $N' = \max(N_\mathrm{T},N_\mathrm{S},N_\mathrm{R})$ and $\sigma_n(\mathbf{H})=\sigma_n(\mathbf{H}_\mathrm{F})=\sigma_n(\mathbf{H}_\mathrm{B})=0$ for $N < n \le N'$.
		We have the following corollaries.

		\begin{corollary}[generic singular value bounds \cite{Fulton2000}]
			\begin{equation}
				\prod_{k \in {K}} \sigma_k(\mathbf{H}) \le \prod_{i \in {I}} \sigma_i(\mathbf{H}_\mathrm{B}) \prod_{j \in {J}} \sigma_j(\mathbf{H}_\mathrm{F}),
				\label{iq:sv_bound_fc}
			\end{equation}
			for all admissible triples $(I, J, K) \in T_r^{N'}$ with $r < N'$, where
			\begin{equation*}
				\begin{gathered}
					% T_r^{N'} \triangleq \Bigl\{(I, J, K) \in U_r^{N'} \mid \text{for all } p < r \text{ and all } (F, G, H) \text{ in } T_p^r,\\
					% \sum_{f \in F} i_f + \sum_{g \in G} j_g \le \sum_{h \in H} k_h + p(p+1)/2 \Bigr\},
					T_r^{N'} \triangleq \Bigl\{(I, J, K) \in U_r^{N'} \mid \forall p < r, (F, G, H) \in T_p^r,\\
					\sum_{f \in F} i_f + \sum_{g \in G} j_g \le \sum_{h \in H} k_h + p(p+1)/2 \Bigr\},
				\end{gathered}
			\end{equation*}
			\begin{equation*}
				U_r^{N'} \triangleq \Bigl\{(I, J, K) \mid \sum_{i \in I} i + \sum_{j \in J} j = \sum_{k \in K} k + r(r+1)/2\Bigr\}.
			\end{equation*}
		\end{corollary}

		\begin{corollary}[upper bound on the largest singular value]
			\begin{equation}
				\sigma_1(\mathbf{H}) \le \sigma_1(\mathbf{H}_\mathrm{B}) \sigma_1(\mathbf{H}_\mathrm{F}).
			\end{equation}
		\end{corollary}

		% \begin{corollary}[lower bound on the smallest singular value]
		% 	\begin{equation}
		% 		\sigma_{N'}(\mathbf{H}) \ge \sigma_{N'}(\mathbf{H}_\mathrm{B}) \sigma_{N'}(\mathbf{H}_\mathrm{F}).
		% 	\end{equation}
		% \end{corollary}

		\begin{corollary}[upper bound on the product of first $k$ singular values]
			\begin{equation}
				\prod_{n=1}^k \sigma_n(\mathbf{H}) \le \prod_{n=1}^k \sigma_n(\mathbf{H}_\mathrm{B}) \prod_{n=1}^k \sigma_n(\mathbf{H}_\mathrm{F}).
			\end{equation}
		\end{corollary}

		% \begin{corollary}[lower bound on the product of last $k$ singular values]
		% 	\begin{equation}
		% 		\prod_{n=N'}^{N'-k+1} \sigma_n(\mathbf{H}) \ge \prod_{n=N'}^{N'-k+1} \sigma_n(\mathbf{H}_\mathrm{B}) \prod_{n=N'}^{N'-k+1} \sigma_n(\mathbf{H}_\mathrm{F}).
		% 	\end{equation}
		% \end{corollary}

		\begin{corollary}[upper bound on the sum of first $k$ singular values to the power of $p$\label{co:sum_power}]
			\begin{equation}
				\sum_{n=1}^k \sigma_n^p(\mathbf{H}) \le \sum_{n=1}^k \sigma_n^p(\mathbf{H}_\mathrm{B}) \sigma_n^p(\mathbf{H}_\mathrm{F}), \quad p > 0.
				\label{iq:sv_bound_fc_power}
			\end{equation}
			When $k = N'$ and $p = 2$, \eqref{iq:sv_bound_fc_power} suggests the channel power is upper bounded by the sum of (sorted) element-wise power product of backward and forward subchannels.
		\end{corollary}

		% \begin{remark}
		% 	From \eqref{eq:scattering_fc} and \eqref{eq:channel_equivalent_fc} in the proof of Proposition \ref{pp:fully_connected}, we notice that \eqref{iq:sv_bound_fc}--\eqref{iq:sv_bound_fc_power} are simultaneously tight when
		% 	% Interestingly, \eqref{iq:sv_bound_fc}--\eqref{iq:sv_bound_fc_power} are simultaneously tight when
		% 	\begin{equation}
		% 		\mathbf{\Theta} = \mathbf{V}_\mathrm{B} \mathbf{U}_\mathrm{F}^\mathsf{H}.
		% 		\label{eq:scattering_fc_tight}
		% 	\end{equation}
		% 	An interpretation is that the off-diagonal entries can enhance the capabilities of
		% 	\begin{itemize}
		% 		\item subspace alignment: $\mathbf{V}_\mathrm{B}$ and $\mathbf{U}_\mathrm{F}^\mathsf{H}$ in \eqref{eq:scattering_fc} fully align the subspaces of $\mathbf{H}_\mathrm{B}$ and $\mathbf{H}_\mathrm{F}$ by rotation;
		% 		\item subchannel rearrangement: $\mathbf{X} = \mathbf{I}$ in \eqref{eq:channel_equivalent_fc} pairs the subchannels of $\mathbf{H}_\mathrm{B}$ and $\mathbf{H}_\mathrm{F}$ from strongest to weakest, which attains the maximum in rearrangement inequality.
		% 	\end{itemize}
		% \end{remark}

		Tight bounds are inapplicable when a \gls{mimo} direct link is present, as the \gls{ris} needs to balance the direct-indirect (additive) and backward-forward (multiplicative) subspace alignments.
		Such a balance often involves optimization approaches and another example will be discussed in Section~\ref{sc:low_complexity}.
	\end{subsection}

\end{section}

\begin{section}{Achievable Rate Maximization}\label{sc:rate}
	The \gls{mimo} achievable rate maximization problem is formulated w.r.t. joint active and passive beamforming
	\begin{maxi!}
		{\scriptstyle{\mathbf{W},\mathbf{\Theta}}}{R = \log \det \biggl(\mathbf{I} + \frac{\mathbf{W}^\mathsf{H}\mathbf{H}^\mathsf{H}\mathbf{H}\mathbf{W}}{\eta}\biggr)}{\label{op:rate}}{\label{ob:rate}}
		\addConstraint{\lVert \mathbf{W} \rVert _\mathrm{F}^2 \le P}{}{}
		\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I}, \quad \forall g,}{}{}
	\end{maxi!}
	where $\mathbf{W}$ is the transmit precoder, $R$ is the achievable rate, $\eta$ is the noise power, and $P$ is the transmit power budget.
	Two methods are proposed below to solve problem \eqref{op:rate}.

	\begin{subsection}{Alternating Optimization}
		Consider an \gls{ao} approach that updates $\mathbf{\Theta}$ and $\mathbf{W}$ iteratively.
		For a given $\mathbf{W}$, the passive beamforming subproblem is
		\begin{maxi!}
			{\scriptstyle{\mathbf{\Theta}}}{\log \det \biggl(\mathbf{I} + \frac{\mathbf{H} \mathbf{Q} \mathbf{H}^\mathsf{H}}{\eta}\biggr)}{\label{op:rate_passive}}{\label{ob:rate_passive}}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I}, \quad \forall g,}{}{}
		\end{maxi!}
		where $\mathbf{Q} \triangleq \mathbf{W} \mathbf{W}^\mathsf{H}$ is the transmit covariance matrix.
		\begin{lemma}\label{lm:rate_gradient}
			The Euclidean gradient of \eqref{ob:rate_passive} w.r.t. \gls{bd}-\gls{ris} block $g$ is
			\begin{equation}
				\frac{\partial R}{\partial \mathbf{\Theta}_g^*} = \frac{1}{\eta} \mathbf{H}_{\mathrm{B},g}^\mathsf{H} \biggl(\mathbf{I} + \frac{\mathbf{H}\mathbf{Q}\mathbf{H}^\mathsf{H}}{\eta}\biggr)^{-1} \mathbf{H} \mathbf{Q} \mathbf{H}_{\mathrm{F},g}^\mathsf{H}.
				\label{eq:rate_gradient}
			\end{equation}
		\end{lemma}

		\begin{proof}
			Please refer to Appendix~\ref{ap:rate_gradient}.
		\end{proof}
		Algorithm \ref{ag:rcg} is then invoked to solve problem \eqref{op:rate} where line \ref{ln:gradient_euclidean} uses \eqref{eq:rate_gradient} explicitly.
		Since \eqref{ob:rate_passive} is a concave function of $\mathbf{\Theta}$, convergence to local-optimal points is guaranteed.
		On the other hand, the global optimal transmit precoder for a fixed $\mathbf{\Theta}$ is given by the eigenmode transmission \cite{Clerckx2013}
		\begin{equation}
			\mathbf{W}^\star = \mathbf{V} {\mathbf{S}^\star}^{1/2},
			\label{eq:precoder_eigenmode}
		\end{equation}
		where $\mathbf{V}$ is the right channel singular matrix and $\mathbf{S}^\star$ is the optimal water-filling power allocation matrix.
		The overall \gls{ao} algorithm converges to local-optimal points of problem \eqref{op:rate} since each subproblem is solved optimally and the objective is bounded above.
	\end{subsection}

	\begin{subsection}{Low-Complexity Solution}\label{sc:low_complexity}
		We then propose a low-complexity solution to problem \eqref{op:rate} based on channel shaping.
		The passive beamforming subproblem \eqref{op:rate_passive} involves transmit covariance matrix $\mathbf{Q}$ and thus requires iterative \gls{rcg} update.
		Instead, we decouple the joint \gls{ris}-transceiver design by recasting \eqref{op:rate_passive} as channel power maximization
		\begin{maxi!}
			{\scriptstyle{\mathbf{\Theta}}}{\lVert \mathbf{H}_\mathrm{D} + \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F} \rVert _\mathrm{F}^2}{\label{op:power_passive}}{\label{ob:power_passive}}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I}, \quad \forall g.}{}{}
		\end{maxi!}
		% which only accounts for channel characteristics but has no trivial solution.
		\begin{remark}
			As mentioned in Section \ref{sc:bounds}, the key of solving \eqref{op:power_passive} is to balance the additive and multiplicative subspace alignments.
			Problem \eqref{op:power_passive} is very similar (in terms of maximizing the inner product of $\mathbf{H}_\mathrm{D}$ and $\mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F}$) to the weighted orthogonal Procrustes problem \cite{Gower2004}
			\begin{mini!}
				{\scriptstyle{\mathbf{\Theta}}}{\lVert \mathbf{H}_\mathrm{D} - \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F} \rVert _\mathrm{F}^2}{\label{op:weighted_orthogonal_procrustes}}{}
				\addConstraint{\mathbf{\Theta}^\mathsf{H} \mathbf{\Theta}=\mathbf{I},}{}{}
			\end{mini!}
			which has no trivial solution.
			One lossy transformation, by moving $\mathbf{\Theta}$ to one side \cite{Bell2003}, formulates standard orthogonal Procrustes problems
			\begin{mini!}
				{\scriptstyle{\mathbf{\Theta}}}{\lVert \mathbf{H}_\mathrm{B}^\dagger \mathbf{H}_\mathrm{D} - \mathbf{\Theta} \mathbf{H}_\mathrm{F} \rVert _\mathrm{F}^2 \text{ or } \lVert \mathbf{H}_\mathrm{D} \mathbf{H}_\mathrm{F}^\dagger - \mathbf{H}_\mathrm{B} \mathbf{\Theta} \rVert _\mathrm{F}^2}{\label{op:standard_orthogonal_procrustes}}{}
				\addConstraint{\mathbf{\Theta}^\mathsf{H} \mathbf{\Theta}=\mathbf{I},}{}{}
			\end{mini!}
			which has global optimal solutions
			\begin{equation}
				\mathbf{\Theta}^\star = \mathbf{U} \mathbf{V}^\mathsf{H}
				\label{eq:orthogonal_procrustes_solution}
			\end{equation}
			where $\mathbf{U}$ and $\mathbf{V}$ are the left and right singular matrices of $\mathbf{H}_\mathrm{B}^\dagger \mathbf{H}_\mathrm{D} \mathbf{H}_\mathrm{F}^\mathsf{H}$ or $\mathbf{H}_\mathrm{B}^\mathsf{H} \mathbf{H}_\mathrm{D} \mathbf{H}_\mathrm{F}^\dagger$ \cite{Golub2013}.
			This suboptimal solution only applies to fully-connected \gls{bd}-\gls{ris}.
		\end{remark}

		Inspired by \cite{Nie2017}, we propose a general solution to problem \eqref{op:power_passive} with arbitrary group size.
		The idea is to successively approximate the quadratic objective \eqref{ob:power_passive} by local Taylor expansions and solve each step in closed form.

		\begin{proposition}\label{pp:power}
			Starting from any $\mathbf{\Theta}^{(0)} \in \mathbb{U}^{N_\mathrm{S} \times N_\mathrm{S}}$, the sequence
			\begin{equation}
				\mathbf{\Theta}_g^{(r+1)} = \mathbf{U}_g^{(r)} \mathbf{V}_g^{(r)}, \quad \forall g.
				\label{eq:scattering_power}
			\end{equation}
			converges to a stationary point of \eqref{op:power_passive}, where $\mathbf{U}_g^{(r)}$ and $\mathbf{V}_g^{(r)}$ are the left and right compact singular matrix of
			\begin{equation}
				\mathbf{M}_g^{(r)} = \mathbf{H}_{\mathrm{B},g}^\mathsf{H} \Bigl(\mathbf{H}_\mathrm{D} + \mathbf{H}_\mathrm{B} \mathrm{diag}\bigl(\mathbf{\Theta}_{[1:g-1]}^{(r+1)},\mathbf{\Theta}_{[g:G]}^{(r)}\bigr) \mathbf{H}_\mathrm{F}\Bigr) \mathbf{H}_{\mathrm{F},g}^\mathsf{H}
				\label{eq:auxiliary_matrix_power}
			\end{equation}
		\end{proposition}

		\begin{proof}
			Please refer to Appendix~\ref{ap:power}.
		\end{proof}
		Once the channel shaping problem \eqref{op:power_passive} is solved, the transmit precoder can be obtained by \eqref{eq:precoder_eigenmode}.
		This two-stage approach decouples both blocks and is computationally efficient.
	\end{subsection}

	% \begin{figure}[!t]
	% 	\centering
	% 	\subfloat[\gls{ris} Elements, $N^\mathrm{T} = 8$\label{fg:pc_rate_sx}]{
	% 		\resizebox{0.48\columnwidth}{!}{
	% 			\input{assets/chapter_5/pc_rate_sx.tex}
	% 		}
	% 	}
	% 	\subfloat[Transmit Antenna, $N^\mathrm{S} = 256$\label{fg:pc_rate_tx}]{
	% 		\resizebox{0.48\columnwidth}{!}{
	% 			\input{assets/chapter_5/pc_rate_tx.tex}
	% 		}
	% 	}
	% 	\caption{Average achievable rate versus group size $L$. $N^\mathrm{R} = 4$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (65, 54, 46) \unit{dB}$.}
	% 	\label{fg:pc_rate}
	% \end{figure}

	% Fig.~\ref{fg:pc_rate_sx} illustrates how \gls{ris} configuration influences the \gls{mimo} \gls{pc} achievable rate.
	% To ensure a \qty{20}{bit/s/Hz} transmission, an \gls{snr} of \qty{13.5}{\dB} is required for a 8T4R system.
	% This value decreases to \qty{12.5}{\dB} (resp. \qty{8}{\dB}) when 32- (resp. 256-) element diagonal \gls{ris} is present.
	% If tetrads can be formed in \gls{bd}-\gls{ris}, the \gls{snr} can be reduced by another \qty{20}{\percent} (resp. \qty{44}{\percent}).
	% Further increase in $L$ yields a marginal gain and incurs $\mathcal{O}(L^2)$ connections.
	% We thus conclude dyadic or tetradic \gls{bd}-\gls{ris} usually strike a good balance between performance and complexity.

	% \begin{figure}[!t]
	% 	\centering
	% 	\resizebox{0.65\columnwidth}{!}{
	% 		\input{assets/chapter_5/power_sx.tex}
	% 	}
	% 	\caption{Average channel power versus \gls{ris} elements $N^\mathrm{S}$ and group size $L$. $(N^\mathrm{T}, N^\mathrm{R}) = (8, 4)$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (65, 54, 46) \unit{dB}$.}
	% 	\label{fg:power_sx}
	% \end{figure}

	% \begin{figure}[!t]
	% 	\centering
	% 	\subfloat[Without Direct Link\label{fg:pc_power_bond_nd}]{
	% 		\resizebox{0.48\columnwidth}{!}{
	% 			\input{assets/chapter_5/pc_power_bond_nd.tex}
	% 		}
	% 	}
	% 	\subfloat[With Direct Link\label{fg:pc_power_bond_hd}]{
	% 		\resizebox{0.48\columnwidth}{!}{
	% 			\input{assets/chapter_5/pc_power_bond_hd.tex}
	% 		}
	% 	}
	% 	\caption{Average channel power versus \gls{ris} group size $L$. $(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R}) = (8, 256, 4)$, $(\Lambda^\mathrm{D}, \Lambda^\mathrm{F}, \Lambda^\mathrm{B}) = (65, 54, 46) \unit{dB}$.}
	% 	\label{fg:pc_power_bond}
	% \end{figure}
	% % Fig.~\ref{fg:power_sx} shows that, apart from numerous reflecting elements, a sufficiently large group size can help
	% % Fig.~\ref{fg:power_sx} shows that increasing the group size can improve the channel power especially for a large \gls{ris}.
	% Fig.~\ref{fg:power_sx} shows that, apart from adding reflecting elements $N^\mathrm{S}$, increasing the group size $L$ also improves the channel power.
	% % The behavior is more obvious for a large \gls{ris}.
	% This behavior is more pronounced for a large \gls{ris}.
	% For example, the gain of pairwise connection is \qty{2.8}{\percent} for $N^\mathrm{S} = 16$ and \qty{28}{\percent} for $N^\mathrm{S} = 256$.
	% % \gls{bd}-\gls{ris} provides a higher channel power than conventional \gls{ris}.
	% % As $L$ increases from 1 to 2,
	% It implies that the channel shaping capability of \gls{bd}-\gls{ris} scales with group size $L$.

	% Fig.~\ref{fg:pc_power_bond_hd} and \ref{fg:pc_power_bond_nd} compare the average channel power without and with direct link.
	% % where the cascaded channel power available to passive \gls{ris} is $\lVert \mathbf{H}^\mathrm{B} \rVert _\mathrm{F}^2 \lVert \mathbf{H}^\mathrm{F} \rVert _\mathrm{F}^2$.
	% % where the cascaded channel power available to passive \gls{ris} is $\lVert \mathbf{H}^\mathrm{B} \rVert _\mathrm{F}^2 \lVert \mathbf{H}^\mathrm{F} \rVert _\mathrm{F}^2$.
	% % ``Cascaded'' means the maximum power of the cascaded channel, i.e., $\lVert \mathbf{H}^\mathrm{B} \rVert _\mathrm{F}^2 \lVert \mathbf{H}^\mathrm{F} \rVert _\mathrm{F}^2$.
	% % maximum power of the cascaded channel, i.e., $\lVert \mathbf{H}^\mathrm{B} \rVert _\mathrm{F}^2 \lVert \mathbf{H}^\mathrm{F} \rVert _\mathrm{F}^2$.
	% % ? ``Cascaded'' means the \emph{power product} of the forward and backward channels.
	% % ``Cascaded'' means the sum of element-wise product of first $N = \min(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R})$ eigenvalues of the forward and backward channels.
	% ``Cascaded'' means the sum of element-wise product of first $N = \min(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R})$ eigenvalues (i.e., element-wise power product) of the forward and backward channels.
	% We observe that diagonal \gls{ris} wastes substantial cascaded power and struggles to align the direct-indirect subspace.
	% When the direct link is absent, only \qty{2.6}{\percent} of available power is utilized by diagonal \gls{ris} while \qty{100}{\percent} power is recycled by fully-connected \gls{ris}.
	% When the direct link is present, the proposed \gls{bd}-\gls{ris} design can balance the direct-indirect and forward-backward subspace alignment for an optimal channel boost.
	% It is worth noting that, when $L$ is sufficiently large, the composite channel power surpasses the power sum of direct and cascaded channels, thanks to the constructive \emph{amplitude superposition} of direct and cascaded channels.
	% This again emphasizes the advantage of in-group connection of \gls{bd}-\gls{ris}.
	% % This is because the direct and indirect channels superpose
	% % the proposed \gls{bd}-\gls{ris} design can also be applied to conventional \gls{ris} to improve the channel power.
	% % When the direct link is present, a small $L$ cannot align the direct-indirect subspace.

	% % neither effectively utilizes the cascaded channel power, nor effectively align the direct-indirect subspace.
	% % We observe that conventional \gls{ris} neither effectively utilizes the cascaded channel power, nor effectively align the direct-indirect subspace.
	% % For conventional \gls{ris},
	% % align the direct-indirect subspace.
	% % When the direct link is present, a small $L$ can neither align the direct-indirect subspace nor the forward-backward subspace.
	% % For a relatively large $N^\mathrm{S}$, conventional \gls{ris} cannot
	% % neither the direct-indirect nor the forward-backward subspace can be effectively aligned with a small $L$.

	% % When the direct link is present, a small $L$ can neither preserve the
	% % neither align the direct-indirect subspace nor the forward-backward subspace.

	% % When the direct link is absent, a large $L$ can align the forward-backward subspace but cannot preserve the forward-backward power balance.

\end{section}

\begin{section}{Simulation Results}\label{sc:simulation}
	In this section, we provide numerical results to evaluate the proposed \gls{bd}-\gls{ris} designs.
	Consider a distance-dependent path loss model $\Lambda(d) = \Lambda_0 d^{-\gamma}$ where $\Lambda_0$ is the reference path loss at distance \qty{1}{m}, $d$ is the propagation distance, and $\gamma$ is the path loss exponent.
	The small-scale fading model is $\mathbf{H} = \sqrt{\kappa/(1+\kappa)} \mathbf{H}_\text{LoS} + \sqrt{1/(1+\kappa)} \mathbf{H}_\text{NLoS}$, where $\kappa$ is the Rician $K$-factor, $\mathbf{H}_\text{LoS}$ is the deterministic \gls{los} component, and $\mathbf{H}_\text{NLoS} \sim \mathcal{CN}(\mathbf{0}, \mathbf{I})$ is the Rayleigh component.
	We set $\Lambda_0=\qty{-30}{dB}$, $d_\mathrm{D}=\qty{14.7}{m}$, $d_\mathrm{F}=\qty{10}{m}$, $d_\mathrm{B}=\qty{6.3}{m}$, $\gamma_\mathrm{D}=3$, $\gamma_\mathrm{F}=2.4$ and $\gamma_\mathrm{B}=2$ for reference, which corresponds to a typical indoor environment with $\Lambda_\mathrm{D}=\qty{-65}{dB}$, $\Lambda_\mathrm{F}=\qty{-54}{dB}$, $\Lambda_\mathrm{B}=\qty{-46}{dB}$.
	The indirect path via \gls{ris} is thus \qty{35}{\dB} weaker than the direct path.
	$\kappa \to \infty$ is assumed for all channels unless otherwise specified.

	\begin{table*}[!t]
		\caption{Average Performance of \gls{bd}-\gls{ris} Designs}
		\label{tb:complexity_test}
		\centering
		\begin{tabular}{ccccccc}
			\toprule
			\multirow{2}{*}{\gls{rcg} path} & \multicolumn{3}{c}{$N_\mathrm{S}=16$} & \multicolumn{3}{c}{$N_\mathrm{S}=256$}                                                                 \\ \cmidrule(lr){2-4} \cmidrule(lr){5-7}
			                                & Objective                              & Iterations                             & Time [s]         & Objective        & Iterations & Time [s]  \\ \midrule
			Geodesic                        & $\num{4.355e-3}$                       & 11.61                                  & $\num{2.038e-2}$ & $\num{1.164e-2}$ & 25.78      & 3.216     \\
			Non-geodesic                    & $\num{4.168e-3}$                       & 169.5                                  & $\num{1.420e-1}$ & $\num{8.873e-3}$ & 278.1      & 27.81     \\ \bottomrule
		\end{tabular}
	\end{table*}

	\begin{subsection}{Channel Singular Values Redistribution}
		\begin{subsubsection}{Pareto Frontier}
			\begin{figure}[!t]
				\centering
				\subfloat[$2 \times 32 \times 2$ (no direct)\label{fg:singular_pareto_sx32_nd}]{
					\resizebox{!}{3.25cm}{
						\input{assets/chapter_5/singular_pareto_sx32_nd.tex}
					}
				}
				\subfloat[$2 \times 32 \times 2$\label{fg:singular_pareto_sx32}]{
					\resizebox{!}{3.25cm}{
						\input{assets/chapter_5/singular_pareto_sx32.tex}
					}
				}
				\\
				\subfloat[$2 \times 64 \times 2$\label{fg:singular_pareto_sx64}]{
					\resizebox{!}{3.25cm}{
						\input{assets/chapter_5/singular_pareto_sx64.tex}
					}
				}
				\subfloat[$2 \times 128 \times 2$\label{fg:singular_pareto_sx128}]{
					\resizebox{!}{3.25cm}{
						\input{assets/chapter_5/singular_pareto_sx128.tex}
					}
				}
				\caption{Pareto frontiers of singular values of a 2T2R channel reshaped by a \gls{ris}.}
				\label{fg:singular_pareto}
			\end{figure}
			% Fig. \ref{fg:singular_pareto} shows the Pareto frontiers of channel singular values reshaped by \gls{bd}-\gls{ris}.
			Fig. \ref{fg:singular_pareto} shows the Pareto singular values of a 2T2R \gls{mimo} reshaped by a \gls{ris}.
			When the direct link is absent, the achievable regions in Fig. \subref*{fg:singular_pareto_sx32_nd} are shaped like pizza slices.
			This is because $\sigma_1(\mathbf{H}) \ge \sigma_2(\mathbf{H}) \ge 0$ and there exists a tradeoff between aligning the two subspaces.
			%  since $\sigma_1(\mathbf{H}) \ge \sigma_2(\mathbf{H}) \ge 0$.
			We observe that the smallest singular value is enhanced up to \num{2e-4} by diagonal \gls{ris} and \num{3e-4} by fully-connected \gls{bd}-\gls{ris}, corresponding to a \qty{50}{\percent} gain.
			When the direct link is present, the shape of the singular value region depends heavily on the relative strength of the indirect link.
			In Fig. \subref*{fg:singular_pareto_sx32}, a 32-element \gls{ris} is insufficient to compensate the \qty{35}{dB} path loss imbalance and results in a limited singular value region that is symmetric around the direct point.
			As the group size $L$ increases, the shape of the region evolves from elliptical to square.
			This transformation not only provides a better tradeoff in subchannel manipulation but also improves the dynamic range of $\sigma_1(\mathbf{H})$ and $\sigma_2(\mathbf{H})$ by \qty{22}{\percent} and \qty{38}{\percent}, respectively.
			The achievable singular value region also enlarges as the number of scattering elements $N_\mathrm{S}$ increases.
			In particular, Fig. \subref*{fg:singular_pareto_sx128} shows that the equivalent channel can be completely nulled by a 128-element \gls{bd}-\gls{ris} but not by a diagonal one.
			Those results demonstrate the superior channel shaping capability of \gls{bd}-\gls{ris} for better signal enhancement and interference suppression.
		\end{subsubsection}

		\begin{subsubsection}{Analytical Bounds and Numerical Results}
			\begin{figure}[!t]
				\centering
				\subfloat[$4 \times 32 \times 4$ (rank-1)\label{fg:singular_bound_rank1_sx32}]{
					\resizebox{!}{3.25cm}{
						\input{assets/chapter_5/singular_bound_rank1_sx32.tex}
					}
				}
				\subfloat[$4 \times 64 \times 4$ (rank-1)\label{fg:singular_bound_rank1_sx64}]{
					\resizebox{!}{3.25cm}{
						\input{assets/chapter_5/singular_bound_rank1_sx64.tex}
					}
				}
				\\
				\subfloat[$4 \times 128 \times 4$ (rank-2)\label{fg:singular_bound_rank2_sx128}]{
					\resizebox{!}{3.25cm}{
						\input{assets/chapter_5/singular_bound_rank2_sx128.tex}
					}
				}
				\subfloat[$4 \times 256 \times 4$ (rank-4)\label{fg:singular_bound_rank4_sx256}]{
					\resizebox{!}{3.25cm}{
						\input{assets/chapter_5/singular_bound_rank4_sx256.tex}
					}
				}
				\caption{
					Achievable channel singular values: analytical bounds (green lines) and numerical optimization results (blue and red bars).
					`D' means diagonal \gls{ris} and `BD' means fully-connected \gls{bd}-\gls{ris}.
					`rank-$k$' refers to the forward channel.
				}
				\label{fg:singular_bound}
			\end{figure}
			Fig. \ref{fg:singular_bound} illustrates the analytical singular value bounds in Proposition \ref{pp:rank_deficient} and the numerical results obtained by solving problem \eqref{op:pareto} with $\rho_n = \pm 1$ and $\rho_{n'} = 0$, $\forall n' \ne n$.
			Here we assme a rank-$k$ forward channel without loss of generality.
			When the \gls{ris} is in the vicinity of the transmitter, Figs. \subref*{fg:singular_bound_rank1_sx32} and \subref*{fg:singular_bound_rank1_sx64} show that the achievable channel singular values indeed satisfy Corollary \ref{co:los}, namely $\sigma_1(\mathbf{H}) \ge \sigma_1(\mathbf{T})$, $\sigma_2(\mathbf{T}) \le \sigma_2(\mathbf{H}) \le \sigma_1(\mathbf{T})$, etc.
			It is obvious that \gls{bd}-\gls{ris} can approach those bounds better than diagonal \gls{ris} especially for a small $N_\mathrm{S}$.
			% Another example is given in Fig. \subref*{fg:singular_bound_rank2_sx128} where the forward channel is rank-2.
			Another example is given in Fig. \subref*{fg:singular_bound_rank2_sx128} with rank-2 forward channel.
			% The observations align with Proposition \ref{pp:rank_deficient} that t
			The first two channel singular values are unbounded above and bounded below by the first two singular values of $\mathbf{T}$, while the last two singular values can be suppressed to zero and bounded above by the first two singular values of $\mathbf{T}$.
			Those observations align with Proposition \ref{pp:rank_deficient} and Corollary \ref{co:extreme}.
			Finally, Fig. \subref*{fg:singular_bound_rank4_sx256} confirms there are no extra singular value bounds when both forward and backward channels are full-rank.
			This can be predicted from \eqref{eq:auxiliary_matrix} where the compact singular matrix $\mathbf{V}_\mathrm{F}$ becomes unitary and $\mathbf{T}=\mathbf{0}$.
			The numerical results are consistent with the analytical bounds, and we conclude that the channel shaping advantage of \gls{bd}-\gls{ris} over diagonal \gls{ris} scales with forward and backward channel ranks.

			\begin{figure}[!t]
				\centering
				\subfloat[$1 \times 256 \times 1$ (no direct)\label{fg:power_bond_txrx1_nd}]{
					\resizebox{!}{3.25cm}{
						\input{assets/chapter_5/power_bond_txrx1_nd.tex}
					}
				}
				\subfloat[$4 \times 256 \times 4$ (no direct)\label{fg:power_bond_txrx4_nd}]{
					\resizebox{!}{3.25cm}{
						\input{assets/chapter_5/power_bond_txrx4_nd.tex}
					}
				}
				\caption{
					Average maximum channel power versus \gls{bd}-\gls{ris} group size and \gls{mimo} dimensions.
					`Cascaded' refers to the available power of the cascaded channel, i.e., the sum of (sorted) element-wise power product of backward and forward subchannels.
				}
				\label{fg:power_bond}
			\end{figure}

			Fig. \ref{fg:power_bond} compares the analytical channel power bound in Corollary \ref{co:sum_power} with $k=N'$, $p=2$ and the numerical results obtained by solving problem \eqref{op:power_passive} when the direct link is absent.
			Here, a fully-connected \gls{bd}-\gls{ris} can attain the upper bound either in closed form or via optimization approach \eqref{eq:scattering_power}.
			% Recall that a fully-connected \gls{bd}-\gls{ris} in closed form \eqref{eq:scattering_fc_tight} can attain the bound.
			% We observe that the bound is also tight for fully-connected \gls{bd}-\gls{ris}.
			% Increasing the group size of \gls{bd}-\gls{ris} significantly improves the channel power especially for a large \gls{mimo}.
			% Apparently, \gls{bd}-\gls{ris} with a larger group size can approach the bound better, and a fully-connected one can achieve the bound.
			% The bound here indicates how much power can potentially be extracted from the forward and backward channels.
			% We observe that using a larger group size can approach the bound better, and full
			% We observe that the bound is tight for fully-connected \gls{bd}-\gls{ris} in \gls{siso} and \gls{mimo}.
			For the \gls{siso} case in Fig. \subref*{fg:power_bond_txrx1_nd}, the maximum channel power is approximately \num{4e-6} by diagonal \gls{ris} and \num{6.5e-6} by fully-connected \gls{bd}-\gls{ris}, corresponding to a \qty{62.5}{\percent} gain.
			This aligns with the asymptotic \gls{bd}-\gls{ris} scaling law derived for \gls{siso} in \cite{Shen2020a}.
			Interestingly, the gain surges to \qty{270}{\percent} in 4T4R \gls{mimo} as shown in Fig. \subref*{fg:power_bond_txrx4_nd}.
			This is because subspace alignment boils down to phase matching in \gls{siso} such that both triangular and Cauchy-Schwarz inequalities in \cite[(50)]{Shen2020a} can be simultaneously tight regardless of the group size.
			That is, diagonal \gls{ris} is sufficient for subspace alignment in \gls{siso} while the \qty{62.5}{\percent} gain from \gls{bd}-\gls{ris} comes purely from subchannel rearrangement (i.e., pairing the forward and backward channels from strongest to weakest).
			Now consider a diagonal \gls{ris} in \gls{mimo}.
			Each element can only apply a common phase shift to the associated rank-1 $N_\mathrm{R} \times N_\mathrm{T}$ indirect channel.
			Therefore, perfect subspace alignment of indirect channels through different elements is generally impossible.
			It means the disadvantage of diagonal \gls{ris} in subspace alignment and subchannel rearrangement scales with \gls{mimo} dimensions.
			We thus conclude that the power gain of \gls{bd}-\gls{ris} scales with group size and \gls{mimo} dimensions.
		\end{subsubsection}

		% The Pareto singular value frontiers of a $2 \times 2$ \gls{mimo} with a 32-element \gls{bd}-\gls{ris} is shown in Fig.~\ref{fg:singular_pareto}.
		% The Pareto frontier of channel singular values reshaped by \gls{bd}-\gls{ris} is shown in Fig. \ref{fg:singular_pareto}.
		% The Pareto frontier of channel singular values reshaped by \gls{bd}-\gls{ris} is shown in Fig. \ref{fg:singular_pareto}.
		% Clearly, a larger group size provides
		% and evolving trend of channel singular values are shown in Fig.~\ref{fg:singular_pareto} and \ref{fg:pc_singular_bound}.
		% Clearly, \gls{bd}-\gls{ris} with a larger group size can redistribute the channel singular values to a wider range.
	\end{subsection}

	\begin{subsection}{Achievable Rate Maximization}
		\begin{figure}[!t]
			\centering
			\subfloat[$16 \times N_\mathrm{S} \times 16$ (no direct)\label{fg:power_sx_txrx16_nd}]{
				\resizebox{!}{3.25cm}{
					\input{assets/chapter_5/power_sx_txrx16_nd.tex}
				}
			}
			\subfloat[$16 \times N_\mathrm{S} \times 16$\label{fg:power_sx_txrx16}]{
				\resizebox{!}{3.25cm}{
					\input{assets/chapter_5/power_sx_txrx16.tex}
				}
			}
			\caption{
				Average maximum channel power versus \gls{ris} configuration.
				`OP-left' and `OP-right' refer to the suboptimal solutions to problem \eqref{op:power_passive} by lossy transformation \eqref{op:standard_orthogonal_procrustes} where $\mathbf{\Theta}$ is to the left and right of the product, respectively.
			}
			\label{fg:power_sx}
		\end{figure}

		We first focus on channel shaping subproblem \eqref{op:power_passive_approx}.
		Fig. \ref{fg:power_sx} shows the achievable channel power under different \gls{ris} configurations.
		An interesting observation is that the relative power gain of \gls{bd}-\gls{ris} over diagonal \gls{ris} is even larger with direct link.
		For example, a 64-element fully \gls{bd}-\gls{ris} can almost provide the same channel power as a 256-element diagonal \gls{ris} in Fig. \ref{fg:power_sx_txrx16}, but not in Fig. \ref{fg:power_sx_txrx16_nd}.
		% This again credits to the superior subspace alignment capability of \gls{bd}-\gls{ris}.
		This is because the \gls{ris} needs to balance the multiplicative forward-backward combining and the additive direct-indirect combining, such that the subspace alignment advantage of \gls{bd}-\gls{ris} becomes more pronounced.
		We also notice that the suboptimal solutions \eqref{eq:orthogonal_procrustes_solution} for fully-connected \gls{bd}-\gls{ris} by lossy transformation \eqref{op:standard_orthogonal_procrustes} are very close to optimal especially for a large $N_\mathrm{S}$.

		\begin{figure}[!t]
			\centering
			\subfloat[$4 \times 128 \times 4$\label{fg:rate_beamforming}]{
				\resizebox{!}{3.25cm}{
					\input{assets/chapter_5/rate_beamforming.tex}
				}
			}
			\subfloat[$N_\mathrm{T} \times 128 \times N_\mathrm{R}$\label{fg:rate_txrx}]{
				\resizebox{!}{3.25cm}{
					\input{assets/chapter_5/rate_txrx.tex}
				}
			}
			\\
			\subfloat[$4 \times N_\mathrm{S} \times 4$\label{fg:rate_sx}]{
				\resizebox{!}{3.25cm}{
					\input{assets/chapter_5/rate_sx.tex}
				}
			}
			\subfloat[$4 \times 128 \times 4$\label{fg:rate_kfactor}]{
				\resizebox{!}{3.25cm}{
					\input{assets/chapter_5/rate_kfactor.tex}
				}
			}
			\caption{
				Average achievable rate versus \gls{mimo} and \gls{ris} configurations.
				The noise power is $\eta = \qty{-75}{dB}$, corresponding to a direct \gls{snr} of \num{-10} to \qty{30}{dB}.
				`Alternate' refers to the alternating optimization and `Decouple' refers to the low-complexity design.
				`D' means diagonal \gls{ris} and `BD' means fully-connected \gls{bd}-\gls{ris}.
			}
			\label{fg:rate}
		\end{figure}

		Fig. \ref{fg:rate} presents the achievable rate under different \gls{mimo} and \gls{ris} configurations.
		At a transmit power of \qty{10}{dB}, Fig. \subref*{fg:rate_beamforming} shows that introducing a 128-element diagonal \gls{ris} to 4T4R \gls{mimo} can improve the achievable rate from \qty{22.2}{bps/Hz} to \qty{29.2}{bps/Hz} ($+\qty{31.5}{\percent}$).
		In contrast, a \gls{bd}-\gls{ris} of group size 4 and 128 can further improve the rate to \qty{32.1}{bps/Hz} ($+\qty{44.6}{\percent}$) and \qty{34}{bps/Hz}  ($+\qty{53.2}{\percent}$), respectively.
		Interestingly, the gap between the optimal \gls{ao} approach \eqref{op:rate_passive}--\eqref{eq:precoder_eigenmode} and the low-complexity solution \eqref{eq:scattering_power} and \eqref{eq:precoder_eigenmode} narrows as the group size increases, and completely vanishes for a fully-connected \gls{bd}-\gls{ris}.
		This implies that the \gls{ris}-transceiver design can be completely decoupled via channel shaping with marginal performance loss.
		Figs. \subref*{fg:rate_txrx} and \subref*{fg:rate_sx} also confirm the advantage of \gls{bd}-\gls{ris} grows with the number of transmit, scatter, and receive antennas.
		In the low power regime (\num{-20} to \qty{-10}{dB}), the slope of the achievable rate is significantly larger with \gls{bd}-\gls{ris}, suggesting that multiple streams can be activated at a much lower \gls{snr}.
		This is because \gls{bd}-\gls{ris} not only spreads the channel singular values to a wider range, but also provides a better tradeoff between subchannels (c.f. Fig. \ref{fg:singular_pareto}).
		Finally, Fig. \subref*{fg:rate_kfactor} shows that the gap between diagonal and \gls{bd}-\gls{ris} narrows as the Rician $K$-factor increases and becomes indistinguishable in \gls{los} environment.
		The observation is expected from previous studies \cite{Shen2020a,Li2023b,Nerini2023} and aligns with Corollary \ref{co:los}, which suggests that the \gls{bd}-\gls{ris} should be deployed in rich-scattering environments to exploit its channel shaping potential.
	\end{subsection}
\end{section}

\begin{section}{Conclusion}
	This paper analyzes the channel shaping capability of \gls{ris} in terms of singular values redistribution.
	We consider a general \gls{bd} architecture that allows elements within the same group to interact, enabling more sophisticated manipulation than diagonal \gls{ris}.
	This translates to a wider dynamic range (with better tradeoff) of singular values and significant power and rate gains, especially in large-scale \gls{mimo} systems.
	We characterize the Pareto frontiers of channel singular values via optimization approach and provide analytical bounds in rank-deficient and fully-connected scenarios.
	An efficient \gls{rcg} algorithm is proposed for smooth \gls{bd}-\gls{ris} optimization problems, which offers lower computation complexity and faster convergence than existing methods.
	We also present two beamforming designs for rate maximization problem, one based on alternating optimization for optimal performance and the other decouples the \gls{ris}-transceiver design for lower complexity.
	Extensive simulations show that the advantage of \gls{bd}-\gls{ris} stems from its superior subspace alignment and subchannel rearrangement capability, which scales with the number of elements, group size, \gls{mimo} dimensions, and channel diversity.

	One future direction is introducing \gls{bd}-\gls{ris} to \gls{mimo} interference channel for interference alignment or cancellation.
	Another open issue is to exploit different groups of \gls{bd}-\gls{ris} to enhance the channel response (and possibly ride extra information) at different frequencies.
	Incorporating a \gls{ris} at both transmitter and receiver sides provides even stronger manipulation that potentially align both direct-indirect and forward-backward subspaces simultaneously.
	% Finally, integrating the \gls{ris} into the \gls{mimo} precoding and combining design is also an interesting topic for future research.
	% into the \gls{mimo} precoding and combining design is also an interesting topic for future research.
	% even better channel shaping can be achieved by incorporating the \gls{ris} into the \gls{mimo} precoding and combining design.
\end{section}