### 2024-03-12 assets/chapter_2/nim_flows.tex
:pencil: Update introduction structure

\begin{tikzpicture}[background rectangle/.style={fill=blue!10}, show background rectangle]

	\def\f{1/exp(((\x)^2)/2)}

	\begin{axis}[thick,ticks=none,domain=-pi:pi,samples=100,axis x line*=middle,axis y line=none,xlabel=\empty,xmin=-4,xmax=4,ymax=4,ymin=-2]

		\addplot[smooth, color=blue] (\x,{sin((9*(deg(x))) )*\f}) coordinate[pos=0.42] (W);
		\addplot[smooth, color=black] (\x,\f) coordinate[pos=0.6] (E);
		\addplot[smooth, color=black] (\x,-\f);

	\end{axis}
	\draw[-latex,thick] (E) to ++(1,0) node[anchor=west] {energy flow};
	\draw[-latex,thick,blue] (W) to ++(-1,0) node[anchor=east,blue] {wave direction};
\end{tikzpicture}