### 2024-03-19 content/chapter_2.tex
:pencil: Add wave scattering models

%!TEX root = ../thesis.tex

\graphicspath{{assets/chapter_2/}}

\chapter{Background and Literature Review}\label{ch:background}

\begin{section}{\glsfmtfull{ris}}
	\begin{subsection}{Programmable Metamaterials}
		Metamaterials refer to artificial structures engineered for unusual properties that may not be found in nature.
		The concept was initially proposed by Victor Veselago in 1967, who conjectured the existence of mediums with negative dielectric constant $\epsilon < 0$ and negative permeability $\mu < 0$ \cite{Veselago1968}.
		Such metamaterials are known as ``negative-index'' because the refraction index is defined as the \emph{negative} square root $n = - \sqrt{\epsilon \mu} < 0$, in order to be consistent with Maxwell's equations.
		It was not until 1999 that their feasibility was experimentally demonstrated by John Pendry at Imperial College using split-ring resonators \cite{Pendry1999}.
		Since then, metamaterials have attracted significant interests due to their counterintuitive properties, to name a few:
		\begin{itemize}
			\item \emph{Negative refraction:} As shown in Fig. \ref{fg:nim_refraction}, the incident and refracted rays stay at the same side of the normal axis \cite{Veselago1968}. This phenomenon is in contrast to the usual refraction but can still be predicted from Snell's law
			\begin{equation}
				\frac{\sin \theta_1}{\sin \theta_2} = n.
			\end{equation}
			It is worth mentioning that a generalized law of refraction and refraction has been proposed in \cite{Yu2011}, which has become a standard reference for the design and analysis of metamaterials.
			\begin{figure}[H]
				\centering
				\subfloat[Negative-Index Material\label{fg:nim_refraction}]{
					\resizebox{0.45\columnwidth}{!}{
						\input{assets/chapter_2/nim_refraction.tex}
					}
				}
				\subfloat[Positive-Index Material\label{fg:pim_refraction}]{
					\resizebox{0.45\columnwidth}{!}{
						\input{assets/chapter_2/pim_refraction.tex}
					}
				}
				\caption{Refraction in negative and positive-index materials. Incident and refracted rays stay at the same side of the normal axis in a negative-index material.}
			\end{figure}
			\item \emph{Opposite wave direction:} As shown in Fig. \ref{fg:nim_flows}, the wave vector and energy flow (indicated by the Poynting vector) are opposite to each other in a negative-index material \cite{Pendry2004}. This can be inferred from the electric field equation
			\begin{equation}
				\vec{E} = \vec{E}_0 \exp(\jmath k z - \jmath \omega t)
			\end{equation}
			where $k = k_0 n < 0$ is the wavenumber, $\vec{E}_0$ and $k_0$ are the free-space electric field and wavenumber reference, $z$ is the propagation distance, $\omega$ is the angular frequency, and $t$ is the time.
			Negative-index materials are thus also called ``left-hand'' because the propagation direction of the electric and magnetic fields can be determined by a left-hand rule.
			\begin{figure}[H]
				\centering
				\subfloat[Negative-Index Material\label{fg:nim_flows}]{
					\resizebox{!}{2cm}{
							\input{assets/chapter_2/nim_flows.tex}
					}
				}
				\subfloat[Positive-Index Material\label{fg:pim_flows}]{
					\resizebox{!}{2cm}{
						\input{assets/chapter_2/pim_flows.tex}
					}
				}
				\caption{Wave and energy have opposite directions in a negative-index material.}
			\end{figure}
		\end{itemize}

		Conventional metamaterials have fixed properties that depend on the geometry and arrangement of their constituent elements.
		Once fabricated, these properties cannot be easily changed unless the structure is physically altered.
		This limits their early usage to military and defense, with applications such as invisibility cloaks and optical illusions.
		In 2014, the concept of ``coding'' and ``programmable'' metamaterials was validated by researchers at Southeast University \cite{Cui2014}, who realized digital control of radar cross-section using biased diodes and \gls{fpga}.
		With a proper model of the target properties and external citation, the metamaterial can be reconfigured in real-time for desired behaviors.
		For example, a self-adaptive metasurface equipped with motion and light sensors have been developed in \cite{Ma2019} for single- and multi-beam steering.

		Next, we discuss the principles of electromagnetic wave redirection via refraction and reflection:
		\begin{itemize}
			\item \emph{Refraction:} As shown in Fig. \ref{fg:nim_focus}, the negative-index material can re-focus the beams diverging from a point source to another point behind the material \cite{Padilla2006}. This could be helpful for wireless applications where the transmitter and receiver are at different sides of the material.
			\item \emph{Reflection:} As shown in Fig. \ref{fg:wave_reflected}, the scattering elements cooperatively alter the phase of the incident wave for a constructive (or destructive) superposition of the reflected waves in the target direction \cite{Poulakis2022}. This could be helpful for wireless applications where the transmitter and receiver are at the same side of the material.
		\end{itemize}
		% electromagnetic wave refraction and reflection.

		\begin{figure}[H]
			\centering
			\subfloat[Negative-Index Material\label{fg:nim_focus}]{
				\resizebox{0.45\columnwidth}{!}{
						\input{assets/chapter_2/nim_focus.tex}
				}
			}
			\subfloat[Positive-Index Material\label{fg:pim_diverge}]{
				\resizebox{0.45\columnwidth}{!}{
					\input{assets/chapter_2/pim_diverge.tex}
				}
			}
			\caption{Refraction through metamaterials. For negative-index material, beams diverging from a point source is set in reverse and converges back to another point.}
		\end{figure}
		\begin{figure}[H]
			\centering
			\subfloat[Incident Wave\label{fg:wave_incident}]{
				\resizebox{!}{4.04cm}{
					\input{assets/chapter_2/wave_incident.tex}
				}
			}
			\subfloat[Reflected Wave\label{fg:wave_reflected}]{
				\resizebox{!}{4cm}{
					\input{assets/chapter_2/wave_reflected.tex}
				}
			}
			\caption{Reflection through metamaterials. Yellow dots represent scattering elements. Solid and dashed lines denote wavefronts and rays, respectively. The scattering elements work together to manipulate the phases of incident waves, resulting in a focused beam steered in the intended direction.}
		\end{figure}
		It is worth noticing that refraction and reflection are different implementations of \emph{passive beamforming} where the phase and amplitude of the ambient signal are altered by the metamaterial for a desired net effect.
		In the next subsection, we will some typical \gls{ris} scattering models and their physical architecture.
	\end{subsection}

	\begin{subsection}{Wave Scattering Models}
		\begin{subsection}{Principles}
			\label{sc:principles}
			\gls{rf} wave can be manipulated by scattering elements made from \emph{programmable metamaterials} or \emph{passive antennas} \cite{Liang2022}.
			As discussed above, the former refracts or reflects the incident signals at the air-cell boundary and mainly applies a phase shift.
			In contrast, the latter allows the wave to feed into and collect from the antenna port such that some energy can be absorbed by the circuit and the rest is reradiated to the space.
			An interesting observation is that when excited by an external wave, a scattering element can simultaneously function as an object and a radiator.
			The corresponding scattered field is \cite{Hansen1989}
			\begin{equation}
				\vec{E}_\text{scatter}(Z_\mathrm{L}) = \underbrace{\vec{E}_\text{structural}}_\text{structural component} + \underbrace{\Gamma I_\mathrm{M} \vec{E}_\text{antenna}}_\text{antenna component},
			\end{equation}
			where $Z_\mathrm{L}$ is the load impedance, $\vec{E}_\text{structural}$ is the residual field under perfect matching (i.e., modelling as an object), $\vec{E}_\text{antenna}$ is the radiated field with unit current at the terminal and no external excitation (i.e., modelling as a radiator), $I_\mathrm{M}$ is the current under perfect matching, and $\Gamma$ is the reflection coefficient
			\begin{equation}
				\Gamma = \frac{Z_\mathrm{L} - Z_0^*}{Z_\mathrm{L} + Z_0},
				\label{eq:reflection_coefficient}
			\end{equation}
			and $Z_0$ is the characteristic impedance for programmable metamaterials or the input impedance for passive antennas.
			It is worth mentioning that
			\begin{itemize}
				\item \emph{Structural component:} Depends on the geometry and material of the scatterer. It is usually modelled as part of the environment multipath \cite{Thomas2012,Liang2020} or simply regarded a \gls{dc} offset when the impinging signal is \gls{cw} \cite{Boyer2014}.
				\item \emph{Antenna component:} Depends on the reflection coefficient that can be altered by load impedance. This is widely exploited for various scattering applications, such as backscatter modulation in \gls{bc} and passive beamforming in \gls{ris} \cite{Zhao2022a}.
			\end{itemize}
			We then introduce two canonical \gls{ris} models that will be adopted in the work chapters.
			More accurate models based on field equations (e.g., \cite{Ozdogan2020,Najafi2020}) and measurement fitting (e.g., \cite{Abeywickrama2020}) are also available in the literature.
			% physics-based (e.g., \cite{Ozdogan2020,Najafi2020}) and practical (e.g., \cite{Abeywickrama2020}) models
		\end{subsection}

		\begin{subsection}{Diagonal Phase Shift Model}
			A straightforward way to model the \gls{ris} scattering effect is to consider independent scattering elements with purely reactive load impedance \cite{Wu2018}.
			The reflection coefficient of the $n$-th element is thus
			\begin{equation}
				\theta_n = \frac{\jmath X_n - Z^*}{\jmath X_n + Z} = \exp(\jmath \phi_n),
			\end{equation}
			where $X_n$ is the reactance and $\phi_n$ is the phase shift on the scattered wave.
			For a total of $N_\mathrm{S}$ elements, the \gls{ris} scattering matrix is \emph{diagonal with complex unit-magnitude entries}
			\begin{equation}
				\mathbf{\Theta} = \mathrm{diag}(\theta_1, \ldots, \theta_{N_\mathrm{S}}) =
				\begin{bmatrix}
					\theta_1 & 0 & \cdots & 0 \\
					0 & \theta_2 & \cdots & 0 \\
					\vdots & \vdots & \ddots & \vdots \\
					0 & 0 & \cdots & \theta_{N_\mathrm{S}}
				\end{bmatrix}.
			\end{equation}
			Despite the strong assumptions, this diagonal phase shift model is widely used for the analysis of \gls{ris} systems due to its simplicity and analytical tractability.
		\end{subsection}

		\begin{subsection}{\glsfmtfull{bd} Model}
			How to model the \gls{ris} response if the passive scattering elements can be cooperative instead of independent?
			This question has been answered by \cite{Shen2020a} where a \gls{bd} model was proposed.
			From a network theory perspective \cite{Ivrlac2010}, the interaction between the scattering elements can be modelled as lossless (but not necessarily symmetric) in-group connections in an $N_\mathrm{S}$-port circuit network, as shown in Figs. \ref{fg:bd_ris_architecture_1} and \ref{fg:bd_ris_architecture_2} \cite{Shen2020a}.
			This architecture allows wave impinging on any element to propagate within the circuit and depart partially from other elements in the same group.
			\begin{figure}[H]
				\centering
				\resizebox{0.75\columnwidth}{!}{
					\includegraphics{assets/chapter_2/bd_ris_architecture_1.eps}
				}
				\caption{Network model of a 4-element \gls{ris} with (a) independent scattering and (b) fully cooperative scattering with all elements interconnected. The figure is from \cite{Shen2020a}.}
				\label{fg:bd_ris_architecture_1}
			\end{figure}
			\begin{figure}[H]
				\centering
				\resizebox{0.75\columnwidth}{!}{
					\includegraphics{assets/chapter_2/bd_ris_architecture_2.eps}
				}
				\caption{Network model of an 8-element \gls{ris} with group-wise cooperative scattering of group size (a) 2 and (b) 4. The group size is a design parameter to balance the circuit complexity and scattering performance. The figure is from \cite{Shen2020a}.}
				\label{fg:bd_ris_architecture_2}
			\end{figure}

			A general \gls{bd}-\gls{ris} can be modeled as an $N_\mathrm{S}$-port network that divides into $G$ individual groups, each containing $L \triangleq N_\mathrm{S} / G$ elements interconnected by real-time reconfigurable connections.
			With symmetric components (e.g., capacitors and inductors), the scattering matrix of group $g \in \mathcal{G} \triangleq \{1, \ldots, G\}$ is \cite{Shen2020a}
			\begin{equation}
				\mathbf{\Theta}_g = (\jmath \mathbf{X}_g + Z_0 \mathbf{I})^{-1} (\jmath \mathbf{X}_g - Z_0 \mathbf{I}),
			\end{equation}
			which satisfies both \emph{symmetric} and \emph{unitary} properties
			\begin{subequations}
				\begin{equation}
					\mathbf{\Theta}_g = \mathbf{\Theta}_g^\mathsf{T},
					\label{eq:symmetric_property}
				\end{equation}
				\begin{equation}
					\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g = \mathbf{I}.
					\label{eq:unitary_property}
				\end{equation}
			\end{subequations}
			On the other hand, lossless networks may also be built over asymmetric passive components (e.g., ring hybrids and branch-line hybrids) \cite{Ahn2006} such that the symmetric constraint \eqref{eq:symmetric_property} can be relaxed.
			This corresponds to the ultimate passive model where energy conservation \eqref{eq:unitary_property} is the only constraint for each group, as widely considered in quantum physics.
			The overall scattering matrix of asymmetric \gls{bd}-\gls{ris} is thus \emph{block-diagonal with unitary blocks}
			\begin{equation}
				\mathbf{\Theta} = \mathrm{diag}(\mathbf{\Theta}_1,\ldots,\mathbf{\Theta}_G) =
				\begin{bmatrix}
					\mathbf{\Theta}_1 & \mathbf{0} & \cdots & \mathbf{0} \\
					\mathbf{0} & \mathbf{\Theta}_2 & \cdots & \mathbf{0} \\
					\vdots & \vdots & \ddots & \vdots \\
					\mathbf{0} & \mathbf{0} & \cdots & \mathbf{\Theta}_G
				\end{bmatrix},
				\label{eq:bd_scattering_matrix}
			\end{equation}
			where \eqref{eq:unitary_property} is equivalently denoted as $\mathbf{\Theta}_g \in \mathbb{U}^{L \times L}$.
			The group size $L$ is a design parameter to balance the circuit complexity and scattering performance.
			Diagonal (single-connected) and unitary (fully-connected) \gls{ris} can be viewed as extreme cases with group size $L=1$ and $L=N_\mathrm{S}$, respectively.
			Therefore, the \gls{bd} model \eqref{eq:bd_scattering_matrix} is envisioned to be the next-generation theoretical foundation for passive \gls{ris}, which grants more design freedom and stronger signal processing capability.

			It is also worth mentioning that each group can be abstracted as a mathematical graph with $L$ vertices and a variable number of edges \cite{Nerini2023a}.
			One element is in the same group with another if and only if there is at least one path (via edges) between them.
			It implies that instead of connecting every pair of elements, the practical circuit can be designed to have a sparse graph with only a few connections, which is beneficial for reducing the circuit complexity and power loss from non-ideal components.
			Antenna directivity and radiation pattern should also be modelled in the scattering matrix, especially when the locations of users or \gls{ris} are not fixed.
			This has motivated the concept of \gls{star}-\gls{ris} \cite{Mu2021,Liu2021} and multi-sector \gls{ris} \cite{Li2023c} where incident wave is partially steered to various directions for different users.

		\end{subsection}

	\end{subsection}

\end{section}

\begin{section}{\glsfmtfull{wpt}}
	\begin{subsection}{Near- and Far-Field Techniques}

	\end{subsection}

	\begin{subsection}{Modules and Coupling Effect}
		\begin{subsubsection}{Block Diagram}

		\end{subsubsection}

		\begin{subsubsection}{Blockwise Coupling}

		\end{subsubsection}
	\end{subsection}

	\begin{subsection}{Non-Linear Harvester Behavior}
		\begin{subsubsection}{Rectifier Circuits}

		\end{subsubsection}

		\begin{subsubsection}{Operation Regions and Signal Models}

		\end{subsubsection}

		\begin{subsubsection}{Impact on Waveform Selection}

		\end{subsubsection}
	\end{subsection}
\end{section}

\begin{section}{\glsfmtfull{swipt}}
	\begin{subsection}{\glsfmtfull{r-e} Tradeoff}

	\end{subsection}

	\begin{subsection}{Receiver Architectures}

	\end{subsection}
\end{section}

\begin{section}{\glsfmtfull{bc}}
	\begin{subsection}{\glsfmtfull{mbc}}

	\end{subsection}

	\begin{subsection}{\glsfmtfull{bbc}}

	\end{subsection}

	\begin{subsection}{\glsfmtfull{ambc}}

	\end{subsection}

	\begin{subsection}{\glsfmtfull{sr}}

	\end{subsection}
\end{section}

% \begin{section}{\glsfmtfull{im}}

% \end{section}

\begin{section}{\glsfmtfull{mimo}}
	\begin{subsection}{\glsfmtfull{pc}: Channel Shaping}

	\end{subsection}

	\begin{subsection}{\glsfmtfull{ic}: Interference Alignment}

	\end{subsection}
\end{section}