### 2024-03-23 assets/chapter_2/receiver_ts.tex

\begin{circuitikz}[transform shape,align=center]
	\draw (0,0) node[bareRXantenna](r){Rx}(r.center)
		to[short] ++(0,-1) node[below]{Time switcher}
		to[short] node[spdt,anchor=in](s){} ++(0.5,0);

	\draw (s.out 1)
		to ++(1.25,0)
		to ++(0,0.5)
		to ++(1,0) node[draw,minimum width=2.5cm,anchor=west](eh){Energy\\harvester};
	\draw (eh.east) to ++(1,0) node[draw,anchor=west,minimum width=2.5cm]{Power\\management};

	\draw (s.out 2)
		to ++(1.25,0)
		to ++(0,-0.5)
		to ++(1,0) node[draw,anchor=west,minimum width=2.5cm]{Information\\decoder};
\end{circuitikz}

% \begin{circuitikz}[transform shape,align=center]
% 	\draw (0,0) node[bareRXantenna](r){Rx}(r.center)
% 		to[short] ++(0,-1)
% 		to node[coupler,anchor=left up](c){Power splitter} ++(1,0);
% 	\draw (c.right up)
% 		to[short] ++(0.5,0)
% 		to ++(0,0.5)
% 		to ++(1,0) node[draw,minimum width=2.5cm,anchor=west](eh){Energy\\harvester};
% 	\draw (eh.east) to ++(1,0) node[draw,anchor=west,minimum width=2.5cm]{Power\\management};
% 	\draw (c.right down)
% 		to[short] ++(0.5,0)
% 		to ++(0,-0.5)
% 		to ++(1,0) node[draw,anchor=west,minimum width=2.5cm]{Information\\decoder};
% \end{circuitikz}