### 2024-03-24 assets/chapter_2/ambc.tex
:pencil: Add abstract and BackCom applications

\begin{tikzpicture}[every node/.style={draw,ultra thick},every path/.style={ultra thick},every text node part/.style={align=center}]
	\node at (0,0) [dart,shape border rotate=90,minimum width=1cm] {};
	\node at (1.5,3.5) [isosceles triangle,isosceles triangle stretches,shape border rotate=+270,minimum size=1cm,minimum height=1cm,anchor=north] {};
	\node at (6.5,3) [rectangle,minimum size=1cm] {};
	\node at (8,0.35) [dart,shape border rotate=270,minimum width=1cm] {};

	\node[draw=none] at (0,-1.375) {Primary\\Transmitter};
	\node[draw=none] at (1,4.375) {Backscatter\\Reader};
	\node[draw=none] at (7,4.375) {Scatter\\Node};
	\node[draw=none] at (8,-1.375) {Primary\\Receiver};

	\draw[blue,decorate,decoration={waves,segment length=4mm,radius=2mm}] (1,-0.5) -- (7.5,-0.5);
	\draw[lightgray,decorate,decoration={waves,segment length=4mm,radius=2mm}] (0.25,1) -- (0.75,3);
	\draw[lightgray,decorate,decoration={waves,segment length=4mm,radius=2mm}] (0.75,0.25) -- (5.5,2.5);
	\draw[lightgray,decorate,decoration={waves,segment length=4mm,radius=2mm}] (5.5,3.5) -- (2.5,3.5);
	\draw[magenta,decorate,decoration={crosses,segment length=16mm,shape size=1.5mm}] (5.11,3.5) -- (2.5,3.5);
	\draw[lightgray,decorate,decoration={waves,segment length=4mm,radius=2mm}] (7.43,3.28) -- (8,1);
	\draw[lightgray,decorate,decoration={crosses,segment length=16mm,shape size=1.5mm}] (7.525,2.9) -- (8,1);

	\node[draw=none] at (4,0.35) {Modulated};
	\node[draw=none] at (4,3.95) {Doubly-};
	\node[draw=none] at (4,3) {Modulated};
\end{tikzpicture}