### 2024-03-24 assets/chapter_2/mbc.tex

\begin{tikzpicture}[every node/.style={draw,ultra thick},every path/.style={ultra thick},every text node part/.style={align=center}]
	\node at (0,0) [kite,minimum size=0.75cm,kite vertex angles=60] {};
	\node at (8,3) [rectangle,minimum size=1cm] {};

	\node[draw=none] at (0,-1.125) {Interrogator};
	\node[draw=none] at (8,4.375) {Scatter\\Node};

	\draw[lightgray,decorate,decoration={snake,post length=2.5mm},-latex] (0.5,0.75) -- (7,3);
	\draw[magenta,decorate,decoration={snake,pre length=2cm,post length=2cm},-latex] (7.25,2.25) -- (0.75,0);

	\node[draw=none] at (3.5,2.625) {Carrier};
	\node[draw=none] at (4.2,0.5) {1};
	\node[draw=none] at (2,-0.125) {0};
	\node[draw=none] at (6.4,1.3) {0};
\end{tikzpicture}