### 2024-03-27 assets/chapter_4/ris.tex
:pencil: Update chapter 4

\begin{tikzpicture}[every node/.style={draw,ultra thick},every path/.style={ultra thick},every text node part/.style={align=center}]
	\node at (0,0) [dart,shape border rotate=90,minimum width=1cm] {};
	\node at (5,3.25) [semicircle,shape border rotate=180,minimum width=1.5cm] {};
	\node at (8,0.35) [dart,shape border rotate=270,minimum width=1cm] {};

	\node[draw=none] at (0,-1) {Transmitter};
	\node[draw=none] at (5,4.4375) {Reflect\\Element};
	\node[draw=none] at (8,-1) {Receiver};

	\draw[blue,decorate,decoration={waves,segment length=4mm,radius=2mm}] (1,-0.25) -- (7.375,-0.25);
	\draw[blue,decorate,decoration={waves,segment length=4mm,radius=2mm}] (0.5,0.75) -- (4,3);
	\draw[blue,decorate,decoration={waves,segment length=4mm,radius=2mm}] (6,3.1625) -- (8,1.25);

	\node[draw=none] at (4.625,1) {Modulated};
\end{tikzpicture}