### 2024-04-30 assets/viva/receiver.tex
:construction: Add viva slides

\begin{tikzpicture}[
		% every axis plot/.append style={thick},
		every node/.append style={draw,minimum width=6cm},
		align=center
	]
	\draw[-{Latex[length=4mm]}] (0,0) node[anchor=south](0){Semi-coherently decode node message} to ++(0,-0.5) node[anchor=north](1){Re-encode for reflection pattern};
	\draw[-{Latex[length=4mm]}] (1.south) to ++(0,-0.5) node[anchor=north](2){Model within equivalent channel};
	\draw[-{Latex[length=4mm]}] (2.south) to ++(0,-0.5) node[anchor=north](3){Coherently decode primary message};
\end{tikzpicture}