### 2024-04-30 assets/viva/wpt_architecture.tex

\begin{circuitikz}[transform shape,align=center]
	\draw(0,0) node[draw](rf){\gls{rf}\\source};
	\draw(rf.east) node[txantenna]{};
	\draw(1,1) node{Tx};
	\draw(4.875,2) node[draw]{\textcolor{orange}{\gls{ris}}};
	\draw(4.875,1) node[draw]{\textcolor{orange}{Channel}};
	\draw(10,0) node[draw](mc){Matching\\network};
	\draw(mc.west) node[rxantenna,xscale=-1]{};
	\draw(8.7,1) node{Rx};
	\draw(mc.east) to ++(0.85,0) node[anchor=west,draw](r){\textcolor{blue}{Rectifier}};
	% \draw(d.east) to ++(0.85,0) node[anchor=west,draw](r){Low-pass\\filter};
	\draw(r.east) to ++(1,0) node[anchor=west,draw](dd){\gls{dc}-\gls{dc}\\converter};
	\draw(dd.east) to ++(1,0) node[anchor=west,draw](s){Storage};

	\draw [dashed] (-0.95,-0.75) rectangle (3.75,2.75);
	\draw (1.4,3) node[]{Energy transmitter};
	\draw [dotted] (6.25,-0.5) rectangle (13.2,2.5);
	\draw (9.9,2) node[]{\textcolor{blue}{Rectenna}};
	\draw [dotted] (13.8,-0.5) rectangle (19.9,1.25);
	\draw (16.75,0.875) node[]{Power management unit};
	\draw [dashed] (6,-0.75) rectangle (20.1,2.75);
	\draw (13.375,3) node[]{Energy receiver};

	\draw (-0.95,-1) node[]{$P_{\mathrm{DC}}^{\mathrm{T}}$};
	\draw (3.75,-1) node[]{$P_{\mathrm{RF}}^{\mathrm{T}}$};
	\draw (6,-1) node[]{$P_{\mathrm{RF}}^{\mathrm{R}}$};
	\draw (13.5,-1) node[]{$P_{\mathrm{DC}}^{\mathrm{R}}$};
	\draw (20.3,-1) node[]{$P_{\mathrm{DC}}^{\mathrm{S}}$};

	\draw[gray,dashdotted,-latex] (10.925,-0.5) to ++(0,-1.5) to ++(-1,0) node[anchor=east,draw](c){Channel\\feedback};
	\draw[gray,dashdotted,latex-] (1.4,-0.75) to ++(0,-1.25) to ++(1,0) node[anchor=west,draw](s){Signal\\optimization};
	\draw[gray,dashdotted,-latex] (c.west) to (s.east);
	\draw[gray] (6.45,-2) node[]{Reverse\\communication link};
\end{circuitikz}