### 2024-05-01 assets/chapter_3/convergence_sca.tex

% This file was created by matlab2tikz.
%
%The latest updates can be retrieved from
%  http://www.mathworks.com/matlabcentral/fileexchange/22022-matlab2tikz-matlab2tikz
%where you can also make suggestions and rate matlab2tikz.
%
\definecolor{mycolor1}{rgb}{0.00000,0.44706,0.74118}%
\definecolor{mycolor2}{rgb}{0.85098,0.32549,0.09804}%
%
\begin{tikzpicture}

\begin{axis}[%
width=3.972in,
height=1.383in,
at={(0.666in,2.433in)},
scale only axis,
xmin=1,
xmax=15,
xtick={ 1,  5, 10, 15},
ymin=23.6,
ymax=24,
ylabel style={font=\color{white!15!black}},
ylabel={DC [$\mu$A]},
axis background/.style={fill=white},
xmajorgrids,
ymajorgrids,
legend style={at={(0.97,0.03)}, anchor=south east, legend cell align=left, align=left, draw=white!15!black},
title style={font=\Large}, label style={font=\Large}, ticklabel style={font=\Large}, legend style={font=\Large}, /pgf/number format/.cd,fixed,precision=4
]
\addplot [color=mycolor1, line width=2.0pt, mark=o, mark options={solid, mycolor1}]
  table[row sep=crcr]{%
1	23.6339898147867\\
2	23.9359168430033\\
3	23.9359212944279\\
4	23.9359212944279\\
5	23.9359212944279\\
6	23.9359212944279\\
7	23.9359212944279\\
8	23.9359212944279\\
9	23.9359212944279\\
10	23.9359212944279\\
11	23.9359212944279\\
12	23.9359212944279\\
13	23.9359212944279\\
14	23.9359212944279\\
15	23.9359212944279\\
};
\addlegendentry{SCA}

\end{axis}

\begin{axis}[%
width=3.972in,
height=1.383in,
at={(0.666in,0.472in)},
scale only axis,
xmin=1,
xmax=15,
xtick={ 1,  5, 10, 15},
xlabel style={font=\color{white!15!black}},
xlabel={Iteration index},
ymin=23.1,
ymax=23.4267428478373,
ylabel style={font=\color{white!15!black}},
ylabel={DC [$\mu$A]},
axis background/.style={fill=white},
xmajorgrids,
ymajorgrids,
legend style={at={(0.97,0.03)}, anchor=south east, legend cell align=left, align=left, draw=white!15!black},
title style={font=\Large}, label style={font=\Large}, ticklabel style={font=\Large}, legend style={font=\Large}, /pgf/number format/.cd,fixed,precision=4
]
\addplot [color=mycolor2, dashed, line width=2.0pt, mark=+, mark options={solid, mycolor2}]
  table[row sep=crcr]{%
1	23.1650475635163\\
2	23.4267359303359\\
3	23.4267428478373\\
4	23.4267428478373\\
5	23.4267428478373\\
6	23.4267428478373\\
7	23.4267428478373\\
8	23.4267428478373\\
9	23.4267428478373\\
10	23.4267428478373\\
11	23.4267428478373\\
12	23.4267428478373\\
13	23.4267428478373\\
14	23.4267428478373\\
15	23.4267428478373\\
};
\addlegendentry{M-SCA}

\end{axis}
\end{tikzpicture}%