### 2024-05-01 assets/chapter_4/energy_distribution.tex

\definecolor{color1}{HTML}{0072BD}
\definecolor{color2}{HTML}{D95319}
\definecolor{color3}{HTML}{EDB120}
\definecolor{color4}{HTML}{7E2F8E}

\begin{tikzpicture}[font=\scriptsize]
	\begin{axis}[
			clip mode=individual,
			height=4cm,
			width=10cm,
			xlabel = $z$,
			ylabel = {Probability Density},
			no markers,
			xmin=0,
			ymin=0,
			xmax=70,
			xtick={0,12.78,20.28,29.62,70},
			xticklabels={$t_0$,$t_1$,$t_2$,$t_3$,$t_4$},
			yticklabels=\empty,
			samples = 200
		]

		\addplot+[very thick,solid,color1] gnuplot[raw gnuplot,smooth] {%
				isint(x) = (int(x)==x);
				gmm(x,rho,lambda)=rho<=0||lambda<=0?1/0:  x<0?0.0:x==0?(rho>1?0.0:rho==1?real(lambda):1/0):  exp(rho*log(lambda)+(rho-1.0)*log(x)-lgamma(rho)-lambda*x);
				set xrange [0:80];
				set yrange [0:1];
				samples=200;
				plot gmm(x,10,1)};
		\addlegendentryexpanded{$f(z \mid x_1)$}

		\addplot+[very thick,dashed,color2] gnuplot[raw gnuplot,smooth] {%
				isint(x) = (int(x)==x);
				gmm(x,rho,lambda)=rho<=0||lambda<=0?1/0:  x<0?0.0:x==0?(rho>1?0.0:rho==1?real(lambda):1/0):  exp(rho*log(lambda)+(rho-1.0)*log(x)-lgamma(rho)-lambda*x);
				set xrange [0:80];
				set yrange [0:1];
				samples=200;
				plot gmm(x,10,0.6)};
		\addlegendentryexpanded{$f(z \mid x_2)$}

		\addplot+[very thick,dotted,color3] gnuplot[raw gnuplot,smooth] {%
				isint(x) = (int(x)==x);
				gmm(x,rho,lambda)=rho<=0||lambda<=0?1/0:  x<0?0.0:x==0?(rho>1?0.0:rho==1?real(lambda):1/0):  exp(rho*log(lambda)+(rho-1.0)*log(x)-lgamma(rho)-lambda*x);
				set xrange [0:80];
				set yrange [0:1];
				samples=200;
				plot gmm(x,10,0.4)};
		\addlegendentryexpanded{$f(z \mid x_3)$}

		\addplot+[very thick,dashdotted,color4] gnuplot[raw gnuplot,smooth] {%
				isint(x) = (int(x)==x);
				gmm(x,rho,lambda)=rho<=0||lambda<=0?1/0:  x<0?0.0:x==0?(rho>1?0.0:rho==1?real(lambda):1/0):  exp(rho*log(lambda)+(rho-1.0)*log(x)-lgamma(rho)-lambda*x);
				set xrange [0:80];
				set yrange [0:1];
				samples=200;
				plot gmm(x,10,0.28)};
		\addlegendentryexpanded{$f(z \mid x_4)$}
		\draw[latex-latex] (0,0) -- (12.78,0) node[above,midway,yshift=-2]{$\mathcal{R}_1$};
		\draw[latex-latex] (12.78,0) -- (20.28,0) node[above,midway,yshift=-2]{$\mathcal{R}_2$};
		\draw[latex-latex] (20.28,0) -- (29.62,0) node[above,midway,yshift=-2]{$\mathcal{R}_3$};
		\draw[latex-latex] (29.62,0) -- (70,0) node[above,midway,yshift=-2]{$\mathcal{R}_4$};
	\end{axis}
\end{tikzpicture}