### 2024-05-01 assets/chapter_4/equivalent_circuit.tex

\begin{circuitikz}[transform shape,font=\large]
	% \tikzstyle{every node}=[font=\Large]
	\draw (0,0) coordinate(O)
	to [sV,l=$V_0$] ++(0,1.5)
	to [L,l=$X_{\text{A}}$] ++(0,1.5)
	to [R=$R_{\text{A}}$,-*] ++(3,0) coordinate(AM)
	to [sI,l_=$I_0$,-*] ++(0,-3)
	to [short] (O);
	\draw (AM)
	to [short] ++(0.5,0)
	to [L,l=$X_m$,-*] ++(3,0) coordinate(M)
	to [R=$R_{\text{S},m}$] ++(3,0) coordinate(MH);
	\draw (M)
	to [R=$R_{\text{P},m}$,-*] ++(0,-3);
	\draw (MH)
	to [D,-*] ++(1.5,0) coordinate(H)
	to [C=$X_{\text{H}}$,-*] ++(0,-3);
	\draw (H)
	to [short] ++(1.5,0)
	to [R=$R_{\text{H}}$] ++(0,-3)
	to [short] (O);

	\draw [dashed] (-1.125,-0.5) rectangle (3.5,4);
	\draw [dashed] (4,-0.5) rectangle (9,4);
	\draw [dashed] (9.5,-0.5) rectangle (13.5,4);

	\draw (1.125,4.25) node[]{Antenna};
	\draw (6.5,4.25) node[]{Modulator};
	\draw (11.5,4.25) node[]{Harvester \& Chip};

	\draw (3.375,2.675) to [short] (3.375,3.175) to [short] (3.25,3.175) to [short,i_=$Z_{\text{A}}$] (3.125,3.175);
	\draw (4.125,-0.125) to [short] (4.125,0.375) to [short] (4.25,0.375) to [short,i=$Z_m$] (4.375,0.375);
	\draw (9.625,-0.125) to [short] (9.625,0.375) to [short] (9.75,0.375) to [short,i=$Z_{\text{H}}$] (9.875,0.375);
\end{circuitikz}