### 2024-05-01 assets/chapter_4/time_structure.tex

\begin{tikzpicture}[font=\footnotesize]
	% MBC/BBC
	\foreach \x in {0}{
			\draw[blue,ultra thick,postaction={pattern=north west lines},pattern color=.] (2*\x,4) rectangle ++(2,0.25);
		}
	\foreach \x in {1}{
			\draw[blue,ultra thick,postaction={pattern=north east lines},pattern color=.] (2*\x,4) rectangle ++(2,0.25);
		}
	\foreach \x in {2}{
			\draw[blue,ultra thick,postaction={pattern=vertical lines},pattern color=.] (2*\x,4) rectangle ++(2,0.25);
		}
	\foreach \x in {3}{
			\draw[blue,ultra thick,postaction={pattern=horizontal lines},pattern color=.] (2*\x,4) rectangle ++(2,0.25);
		}
	\draw[latex-latex] (0,4.35) -- (2,4.35) node[above,midway]{BB};
	\node at (-1.5,4.125) {MBC/BBC};

	% AmBC/SR
	\foreach \x in {0,...,7}{
			\draw[teal,fill=teal!35!white,postaction={pattern=north west lines},pattern color=.] (0.5*\x,3) rectangle ++(0.5,0.25);
		}
	\foreach \x in {8,...,15}{
			\draw[teal,fill=teal!35!white,postaction={pattern=north east lines},pattern color=.] (0.5*\x,3) rectangle ++(0.5,0.25);
		}
	\foreach \x in {0,...,1}{
			\draw[teal,ultra thick] (4*\x,3) rectangle ++(4,0.25);
		}
	\draw[latex-latex] (0,3.35) -- (4,3.35) node[above,midway]{BB};
	\draw[latex-latex] (0,2.9) -- (0.5,2.9) node[below,midway]{PB};
	\node at (-1.5,3.125) {AmBC/SR};

	% RIS
	\foreach \x in {0,...,15}{
			\draw[red,fill=red!35!white,postaction={pattern=north east lines},pattern color=.] (0.5*\x,2) rectangle ++(0.5,0.25);
		}
	\draw[red,ultra thick] (0,2) rectangle ++(8,0.25);
	\draw[latex-latex] (0,2.35) -- (8,2.35) node[above,midway]{CB};
	\draw[latex-latex] (0,1.9) -- (0.5,1.9) node[below,midway]{PB};
	\node at (-1.5,2.125) {RIS};

	% RIScatter
	\foreach \x in {0,...,3}{
			\draw[brown,fill=brown!35!white,postaction={pattern=north west lines},pattern color=.] (0.5*\x,1) rectangle ++(0.5,0.25);
		}
	\foreach \x in {4,...,15}{
			\draw[brown,fill=brown!35!white,postaction={pattern=north east lines},pattern color=.] (0.5*\x,1) rectangle ++(0.5,0.25);
		}
	\foreach \x in {0,...,3}{
			\draw[brown,ultra thick] (2*\x,1) rectangle ++(2,0.25);
		}
	\draw[latex-latex] (0,1.35) -- (2,1.35) node[above,midway]{BB};
	\draw[latex-latex] (0,0.9) -- (0.5,0.9) node[below,midway]{PB};

	\node at (-1.5,1.125) {RIScatter};

	\draw [pattern=north west lines,ultra thick](0.15,5) rectangle ++(0.5,0.25) node[xshift=20,yshift=-3.5] {State 1};
	\draw [pattern=north east lines,ultra thick](2.15,5) rectangle ++(0.5,0.25) node[xshift=20,yshift=-3.5] {State 2};
	\draw [pattern=vertical lines,ultra thick](4.15,5) rectangle ++(0.5,0.25) node[xshift=20,yshift=-3.5] {State 3};
	\draw [pattern=horizontal lines,ultra thick](6.15,5) rectangle ++(0.5,0.25) node[xshift=20,yshift=-3.5] {State 4};
\end{tikzpicture}