### 2024-05-01 assets/viva/waveform_subband.tex

% This file was created by matlab2tikz.
%
%The latest updates can be retrieved from
%  http://www.mathworks.com/matlabcentral/fileexchange/22022-matlab2tikz-matlab2tikz
%where you can also make suggestions and rate matlab2tikz.
%
\definecolor{mycolor1}{rgb}{0.00000,0.44700,0.74100}%
\definecolor{mycolor2}{rgb}{0.85000,0.32500,0.09800}%
%
\begin{tikzpicture}

\begin{axis}[%
width=4.036in,
height=0.462in,
at={(0.677in,3.21in)},
scale only axis,
xmin=0,
xmax=2,
xtick={1},
ymin=0,
axis background/.style={fill=white},
xmajorgrids,
ymajorgrids,
legend style={at={(0.5,1.03)}, anchor=south, legend columns=2, legend cell align=left, align=left, draw=white!15!black},
title style={font=\Large}, label style={font=\Large}, ticklabel style={font=\Large}, legend style={font=\Large}
]
\addplot[ycomb, color=mycolor1, line width=2.0pt, mark=o, mark options={solid, mycolor1}] table[row sep=crcr] {%
1	4.47213593809508\\
};
\addplot[forget plot, color=white!15!black, line width=2.0pt] table[row sep=crcr] {%
0	0\\
2	0\\
};
\addlegendentry{{\boldmath${s}$}$_{\mathrm{I}}$}

\addplot[ycomb, color=mycolor2, line width=2.0pt, mark=x, mark options={solid, mycolor2}] table[row sep=crcr] {%
1	0.000358590327058133\\
};
\addplot[forget plot, color=white!15!black, line width=2.0pt] table[row sep=crcr] {%
0	0\\
2	0\\
};
\addlegendentry{{\boldmath${s}$}$_{\mathrm{P}}$}

\end{axis}

\begin{axis}[%
width=4.036in,
height=0.462in,
at={(0.677in,2.522in)},
scale only axis,
xmin=0,
xmax=3,
xtick={1, 2},
ymin=0,
axis background/.style={fill=white},
xmajorgrids,
ymajorgrids,
title style={font=\Large}, label style={font=\Large}, ticklabel style={font=\Large}, legend style={font=\Large}
]
\addplot[ycomb, color=mycolor1, line width=2.0pt, mark=o, mark options={solid, mycolor1}, forget plot] table[row sep=crcr] {%
1	0.550823184768846\\
2	4.36837845097092\\
};
\addplot[forget plot, color=white!15!black, line width=2.0pt] table[row sep=crcr] {%
0	0\\
3	0\\
};
\addplot[ycomb, color=mycolor2, line width=2.0pt, mark=x, mark options={solid, mycolor2}, forget plot] table[row sep=crcr] {%
1	1.48597207854411e-05\\
2	1.53805823004565e-05\\
};
\addplot[forget plot, color=white!15!black, line width=2.0pt] table[row sep=crcr] {%
0	0\\
3	0\\
};
\end{axis}

\begin{axis}[%
width=4.036in,
height=0.462in,
at={(0.677in,1.834in)},
scale only axis,
xmin=0,
xmax=5,
xtick={1, 2, 3, 4},
ymin=0,
ylabel style={font=\color{white!15!black}},
ylabel={Waveform amplitude},
axis background/.style={fill=white},
xmajorgrids,
ymajorgrids,
title style={font=\Large}, label style={font=\Large}, ticklabel style={font=\Large}, legend style={font=\Large}
]
\addplot[ycomb, color=mycolor1, line width=2.0pt, mark=o, mark options={solid, mycolor1}, forget plot] table[row sep=crcr] {%
1	0.0326721056203988\\
2	0.0697944085211443\\
3	0.193520763120912\\
4	1.91750698481035\\
};
\addplot[forget plot, color=white!15!black, line width=2.0pt] table[row sep=crcr] {%
0	0\\
5	0\\
};
\addplot[ycomb, color=mycolor2, line width=2.0pt, mark=x, mark options={solid, mycolor2}, forget plot] table[row sep=crcr] {%
1	0.972503011468572\\
2	1.22779327868874\\
3	1.48063097057209\\
4	2.13162879975205\\
};
\addplot[forget plot, color=white!15!black, line width=2.0pt] table[row sep=crcr] {%
0	0\\
5	0\\
};
\end{axis}

\begin{axis}[%
width=4.036in,
height=0.462in,
at={(0.677in,1.146in)},
scale only axis,
xmin=0,
xmax=9,
xtick={1, 2, 3, 4, 5, 6, 7, 8},
ymin=0,
ytick={0, 1},
axis background/.style={fill=white},
xmajorgrids,
ymajorgrids,
title style={font=\Large}, label style={font=\Large}, ticklabel style={font=\Large}, legend style={font=\Large}
]
\addplot[ycomb, color=mycolor1, line width=2.0pt, mark=o, mark options={solid, mycolor1}, forget plot] table[row sep=crcr] {%
1	0.00676327571402152\\
2	0.00930781551713654\\
3	0.0129863266085796\\
4	0.0182780278377859\\
5	0.0264785684252956\\
6	0.0417816477229242\\
7	0.0921547943904462\\
8	0.387871897134795\\
};
\addplot[forget plot, color=white!15!black, line width=2.0pt] table[row sep=crcr] {%
0	0\\
9	0\\
};
\addplot[ycomb, color=mycolor2, line width=2.0pt, mark=x, mark options={solid, mycolor2}, forget plot] table[row sep=crcr] {%
1	0.897080828535262\\
2	1.11267484124012\\
3	1.2760937466216\\
4	1.42973823824102\\
5	1.54367522426452\\
6	1.64897920494507\\
7	1.73956386270218\\
8	1.89059947472085\\
};
\addplot[forget plot, color=white!15!black, line width=2.0pt] table[row sep=crcr] {%
0	0\\
9	0\\
};
\end{axis}

\begin{axis}[%
width=4.036in,
height=0.462in,
at={(0.677in,0.458in)},
scale only axis,
xmin=0,
xmax=17,
xtick={ 1,  2,  3,  4,  5,  6,  7,  8,  9, 10, 11, 12, 13, 14, 15, 16},
xlabel style={font=\color{white!15!black}},
xlabel={Sorted subband index},
ymin=0,
ytick={0, 1},
axis background/.style={fill=white},
xmajorgrids,
ymajorgrids,
title style={font=\Large}, label style={font=\Large}, ticklabel style={font=\Large}, legend style={font=\Large}
]
\addplot[ycomb, color=mycolor1, line width=2.0pt, mark=o, mark options={solid, mycolor1}, forget plot] table[row sep=crcr] {%
1	0.00205132551819701\\
2	0.00235447373497551\\
3	0.00272196136127332\\
4	0.00316588966668174\\
5	0.00369517036368117\\
6	0.00432537450541609\\
7	0.00508695652011363\\
8	0.00600094215308655\\
9	0.00710012267822671\\
10	0.00851262778314969\\
11	0.0103510448032954\\
12	0.0128569172775466\\
13	0.0164283761517276\\
14	0.0229541038226562\\
15	0.0402894292810385\\
16	0.0926842285796484\\
};
\addplot[forget plot, color=white!15!black, line width=2.0pt] table[row sep=crcr] {%
0	0\\
17	0\\
};
\addplot[ycomb, color=mycolor2, line width=2.0pt, mark=x, mark options={solid, mycolor2}, forget plot] table[row sep=crcr] {%
1	0.628841069483884\\
2	0.717675502523908\\
3	0.795496148359302\\
4	0.864626670343281\\
5	0.930591126638763\\
6	0.989724762400551\\
7	1.04508455096183\\
8	1.09678731061405\\
9	1.14380160245268\\
10	1.18636001210616\\
11	1.2248217690153\\
12	1.2585853066132\\
13	1.28750705323146\\
14	1.31224542931104\\
15	1.33460858167442\\
16	1.35822969320702\\
};
\addplot[forget plot, color=white!15!black, line width=2.0pt] table[row sep=crcr] {%
0	0\\
17	0\\
};
\end{axis}
\end{tikzpicture}%