### 2024-05-01 content/abstract.tex

% ************************** Thesis Abstract *****************************
% Use `abstract' as an option in the document class to print only the titlepage and the abstract.
\begin{abstract}
	\gls{ris} is one of the most promising physical-layer technology for 6G.
	It adopts numerous low-power scattering elements to customize the propagation environment for improved network performance, unlocking a new potential of \emph{channel design} as never before.
	In this dissertation, we first provide an overview of its physical structure, characteristics, applications, and scattering models, then discuss how it addresses the key issues in \gls{swipt} and compare it with \gls{bc} in terms of functionalities and preferences.
	The work chapters investigate three typical use cases of \gls{ris}: Joint design with transceiver for a specific performance measure (beamforming); ride its own information over legacy networks (modulation); and manipulate the inherent properties of the wireless environment as a stand-alone device (channel shaping).
	In particular, we address the following topics:
	\begin{itemize}
		\item \emph{\gls{ris}-aided \gls{swipt}:} We introduce \gls{ris} into multi-antenna, multi-carrier \gls{swipt} systems, investigating joint waveform and beamforming design. \gls{r-e} region of practical receivers are characterized under non-linear harvester and frequency-flat \gls{ris} models.
		\item \emph{RIScatter:} This novel scatter protocol integrates \gls{ris} and \gls{bc} from an input distribution perspective. The reflection pattern is exploited simultaneously for beamforming of the primary link and modulation of the backscatter link. We propose a practical cooperative receiver and characterize the achievable rate region.
		\item \emph{\gls{mimo} channel shaping:} We exploit an advanced \gls{bd}-\gls{ris} architecture for singular value redistribution in \gls{pc} and leakage interference minimization in \gls{ic}.
		Their implications on joint \gls{ris}-transceiver designs are also discussed.
		We also propose an efficient and universal \gls{bd}-\gls{ris} design framework.
	\end{itemize}
\end{abstract}